---
title:  "vaira-ensemble-postprocessing"
author: "Xiulin Gao"
email:  "xiulingao@lbl.gov"
date:   "15-Aug-2022"
---

## Introduction

This script is used to process ensemble runs at Vaira Ranch, California. We want to do 1. global parameter sensitivity analysis; 2. filter survived ensembles by comparing model output to site observations. This scripts uses a set of scripts developed by Marcos Longo at <https://github.com/mpaiao/FATES_Utils>; part of this script is also adapted from Marcos Longo's postp-rocessing script, significant changes are made to do post-processing specifically for 1PFT ensemble runs at site.

```{r, label = 'reset-R',message=FALSE, results='hide'}
# Unload all packages except for the R default ones
plist = names(sessionInfo()$otherPkgs)
if (length(plist) > 0){
   dummy = sapply(X=paste0("package:",plist),FUN=detach,character.only=TRUE,unload=TRUE)
}#end if (length(plist) > 0)


# Remove all variables
rm(list=ls())

# Reset warnings
options(warn=0)

# Close all plots
invisible(graphics.off())

# Clean up
invisible(gc())
```




```{r, label="path"}
home_path   = path.expand("~")
main_path   = file.path("/Volumes/Ldian/FATES/sierra-refine0716/fireoff")
#fireon_path = file.path(main_path,"fireon-site/2022-11-02")
util_path   = file.path(home_path,"Util/RUtils")

ens_cases  = c("cz2-p61-fireoff-tuneFir.nc",
               "tha-p61-fireoff-tuneFir.nc"
               )

case_files  = file.path(main_path,ens_cases)
ncase       = length(ens_cases)
case_names  = c(1L,2L)

case_desc   =   c(
                  "p61-refine-cz2",
                  "p61-refine-tha"
                  )

file_desc   = "sierra-refine-2sites-0716"

names(case_desc) = c("1","2")


#obs path
#obs_main    = file.path("~/Google Drive/My Drive/CA-OakWoodland/benchmark")
#flux_path   = file.path(obs_main,"tonzi-gpp-energy-monthmean.csv")
#mas_path    = file.path(obs_main,"WRF_CA-HarwoodBiomass.csv")
#can_path    = file.path(obs_main,"WRF_CA-HarwoodCanopyCover.csv")
#ras_main    = file.path("~/Google Drive/My Drive/CA-grassland-simulationDoc/benchmark/CA-GPP-LAI-005degree/")
#lai_path    = file.path(ras_main,"LAI-onWRFgrids.csv")
#gpp_path    = file.path(ras_main,"GPP-onWRFgrids.csv")
#frap_raw    = file.path("~/Google Drive/My Drive/CA-grassland-simulationDoc/benchmark/CA-grassfire-climate/fire20_1.gdb")


```




```{r,label="make-dir"}

#plot path
plot_main  = file.path(main_path,"Figures")
# Output path for time series
tsage_path    = file.path(plot_main,"tseries_age"         )
tsdbh_path    = file.path(plot_main,"tseries_dbh"         )
tspft_path    = file.path(plot_main,"tseries_pft"         )
tsmort_path   = file.path(plot_main,"tseries_mort"        )
tstheme_path  = file.path(plot_main,"tseries_theme"       )
secy_path     = file.path(plot_main,"seasonal-cycle"      )
sens_path     = file.path(plot_main,"variance-decomp"     )

# Create paths for time series
dummy = dir.create(tsage_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsdbh_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tspft_path  , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tsmort_path , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(tstheme_path, recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(secy_path   , recursive = TRUE, showWarnings = FALSE)
dummy = dir.create(sens_path   , recursive = TRUE, showWarnings = FALSE)

```




```{r,label="user-pft-fuel"}
user_pftinfo  = TRUE
user_fuelinfo = TRUE

# List of PFTs, in case user_pftinfo is TRUE
n            = 0
pftinfo      = list()
n            = n + 1
pftinfo[[n]] = list( id     = 1
                   , key    = "pft1"
                   , short  = "Pine"
                   , desc   = "Pine tree"
                   , colour = "#74A089"
                   , parse  ="P*i*n*e[E*v*g*r*n]"
                   )
n            = n + 1
pftinfo[[n]] = list( id     = 2
                   , key    = "pft2"
                   , short  = "Cedar"
                   , desc   = "Incense cedar"
                   , colour = "#9A8822"
                   , parse  ="C*e*d*a*r[E*v*g*r*n]"
                   )       
n            = n + 1
pftinfo[[n]] = list( id     = 3
                   , key    = "pft3"
                   , short  = "Fir"
                   , desc   = "White fir"
                   , colour = "#F8AFA8"
                   , parse  ="F*i*r[E*v*g*r*n]"
                   ) 
n            = n + 1
pftinfo[[n]] = list( id     = 4
                   , key    = "pft4"
                   , short  = "Oak"
                   , desc   = "Black oak"
                   , colour = "#FDDDA0"
                   , parse  ="F*i*r[E*v*g*r*n]"
                   ) #end of PFT list

# List of fuels, in case user_fuleinfo is TRUE
n               = 0
fuelinfo        = list()
n               = n + 1
fuelinfo[[n]]   = list( id      = 1
                      , key     = "fuel1"
                      , short   = "Twig"
                      , desc    = "Fuel from twig" 
                      , colour  = "#E69F00"
                      , parse   = "F*u*e*l[T*w*i*g]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 2
                      , key     = "fuel2"
                      , short   = "SBranch"
                      , desc    = "Fuel from small branch" 
                      , colour  = "#56B4E9"
                      , parse   = "F*u*e*l[S*b*r*a*n*c*h]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 3
                      , key     = "fuel3"
                      , short   = "LBranch"
                      , desc    = "Fuel from large branch" 
                      , colour  = "#0072B2"
                      , parse   = "F*u*e*l[L*b*r*a*n*c*h]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 4
                      , key     = "fuel4"
                      , short   = "Trunk"
                      , desc    = "Fuel from trunk" 
                      , colour  = "#D55E00"
                      , parse   = "F*u*e*l[T*r*u*n*k]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 5
                      , key     = "fuel5"
                      , short   = "DeadLeaf"
                      , desc    = "Fuel from dead leaf" 
                      , colour  = "#999999"
                      , parse   = "F*u*e*l[D*l*e*a*f]")
n               = n + 1
fuelinfo[[n]]   = list( id      = 6
                      , key     = "fuel6"
                      , short   = "LiveGrass"
                      , desc    = "Fuel from live grass" 
                      , colour  = "#009E73"
                      , parse   = "F*u*e*l[L*g*r*a*s*s]") #end of fuel list



```





```{r,label="plot-setting"}

gg_device  = c("png")     # Output devices to use (Check ggsave for acceptable formats)
gg_depth   = 600          # Plot resolution (dpi)
gg_ptsz    = 18           # Font size
gg_ptszl   = 26
gg_width   = 17.5         # Plot width (units below)
gg_widthn  = 14.5
gg_height  = 8.5          # Plot height (units below)
gg_units   = "in"         # Units for plot size
gg_screen  = FALSE         # Show plots on screen as well?
gg_tfmt    = "%Y"         # Format for time strings in the time series %Y: 4 DIGITS YEAR %y 2 digits year
gg_ncolours    = 129      # Number of node colours for heat maps.
gg_fleg        = 1./6.    # Fraction of plotting area dedicated for legend
gg_dstat_thick = 0.1      # Thickess for the drought-deciduous status band in the stress overview plot (Value


ens_colour = c("#A0AE6A", "#836B43", "#D68D18", "#437683", "#18B0D6")

top12 <- c("#88CCEE", "#CC6677", "#DDCC77", "#117733", "#332288", "#AA4499", 
           "#44AA99", "#999933", "#882255", "#6699CC", "#888888", "#661100")

site_colour  = "#000000"

ndevice = length(gg_device)


# Define lower and upper bound for interannual variability ribbon around the mean.
sdev_ribbon  = 1.                  # Standard-deviation equivalent for ribbon 
qlwr_ribbon  = pnorm(-sdev_ribbon) # Lower quantile for ribbon
qupr_ribbon  = pnorm(+sdev_ribbon) # Lower quantile for ribbon
alpha_ribbon = 0.2

```



```{r, label="load-everything"}
source(file.path(util_path,"load.everything.r"),chdir=TRUE)

```



```{r,label="ensemble-label"}

ninst     =72
n_task    =36
n_node    =ceiling(ninst/n_task)
n_cpuls   = n_task-round((n_node-(ninst/n_task))*n_task,digits=0)
cpuls     = seq(n_cpuls)
task_ens = rep(seq(n_task), times = n_node-1)
task_ens = append(task_ens,cpuls)
node_ens = rep(seq(n_node),each=n_task)
len_node = length(node_ens)
#node_ens = node_ens[-c(1501:len_node)]
nodekeys = sprintf("node_%3.3i",node_ens)
taskeys  = sprintf("task_%3.3i",task_ens)
ens_labels = paste0(nodekeys,"-",taskeys)

```






```{r,label="time-setting"}

nc_conn      = nc_open(case_files[1])
time_val     = ncvar_get(nc_conn,"time")
s_t          = 1
e_t          = 2400
lst_s        = 1779
lst_e        = 1799
start_y      = 1600
end_y        = 1799
time_use     = time_val[s_t:e_t]
simu_time    = c(start_y:end_y)
model_yra    = 1600
n_yr         = round((tail(time_use,1)-head(time_use,1))/365,digits=0)
yr_start     = model_yra + floor(time_use[1]/365)
yr_end       = yr_start + n_yr-1
nc_year      = as.numeric(yr_start:yr_end)
nc_month     = as.numeric(1:12)
yr_each      = rep(nc_year,each=12)
yr_ens       = rep(yr_each,times=ninst)
month_each   = rep(nc_month, times=n_yr)
month_ens    = rep(month_each, times=ninst)
ens_tstamp   = make_datetime(year=yr_ens,month=month_ens,day=1L)
ntstamp_inst = length(ens_tstamp)
each_tstamp  = make_datetime(year=yr_each,month=month_each,day=1L)
ntstamp_each = length(each_tstamp)
dummy        = nc_close(nc_conn)

```




```{r,label='redefine-read-group',message=FALSE,results='hide'}

read_age   = TRUE
read_dbh   = TRUE
read_fuel  = TRUE
read_hlm2d = FALSE
read_phen  = FALSE
read_soi   = FALSE

read_var  = c( "fates_basalarea","fates_crownarea",	
               "fates_canopycrownarea","fates_ddbh_canopy",
               "fates_ddbh_ustory","fates_ddbh","fates_fuel_amount",
               "fates_gpp","fates_npp","fates_lai_canopy","fates_lai_ustory",
               "fates_lai","fates_mortality_background","fates_mortality_cstarv",
               "fates_mortality_freezing","fates_mortality_fire","fates_mortality_ustory",
               "fates_mortality_hydraulic","fates_mortality_termination",
               "fates_nplant","fates_nplant_canopy","fates_nplant_ustory",
               "fates_vegc_aboveground",
               "fates_mortality_rxfire",
               "fates_mortality_rxcambial","fates_mortality_cambialburn",
               "fates_mortality_crownscorch","fates_mortality_rxcrown",
               "fates_seed_alloc") 

fatesvar_new  = fatesvar                   %>% 
                filter(vnam %in% read_var)
nfatesvar_new = length(fatesvar_new$vnam)

### define variable to read for size class only
szpf_var = c("fates_nplant","fates_basalarea",
             "fates_nplant_canopy","fates_nplant_ustory",
             "fates_mortality_fire","fates_mortality_rxfire",
             "fates_mortality_ustory",
             "fates_ddbh_canopy","fates_ddbh_ustory","fates_ddbh",
             "fates_mortality_termination")
fatesvar_szpf = fatesvar     %>%  filter(vnam %in% szpf_var) 
nszpf_read    = length(fatesvar_szpf$vnam)

age_var  = c("fates_vegc","fates_patcharea","fates_fuel_amount",
             "fates_rxfire_burnfrac","fates_burnfrac")
fatesvar_ap  = fatesvar %>% filter(vnam %in% age_var)
nage_read    = length(fatesvar_ap$vnam)
               

```




```{r,label="dimension-setting"}
if ("nc_conn" %in% ls()){dummy = nc_close(nc_conn); rm(nc_conn)}

nc_dlist <- list()
nc_vlist <- list()

nc_conn  = nc_open(filename=case_files[1])
nc_nvars = nc_conn$nvars
nc_ndims = nc_conn$ndims
nc_dlist = rep(NA_character_,times=nc_ndims)
nc_vlist = rep(NA_character_,times=nc_nvars)
for (d in sequence(nc_ndims)) nc_dlist[d] = nc_conn$dim[[d]]$name
for (v in sequence(nc_nvars)) nc_vlist[v] = nc_conn$var[[v]]$name

#---~---
# Gather dimension information, then initialize matrices
#---~---

# List of age classes
if(read_age){
idxage   = match("fates_levage",nc_dlist)
if (is.finite(idxage)){
   ages     = nc_conn$dim[[idxage]]$vals
   nages    = nc_conn$dim[[idxage]]$len

   ageinfo  = tibble( id      = sequence(nages)
                    , age_lwr = ages
                    , age_upr = c(ages[-1],Inf)
                    , key     = sprintf("age_%3.3i",ages)
                    , desc    = c( paste0("paste(paste(",age_lwr[-nages],"<= A*g*e)<",age_upr[-nages],"*y*r)")
                                 , paste0("paste( A*g*e >=",age_upr[nages],"*y*r)")
                                 )#end c
                    , labs    = c( paste0("paste(",age_lwr[-nages],"-",age_upr[-nages],")")
                                 , paste0("paste(",age_lwr[ nages],"-infinity)")
                                 )#end c
                    , colour  = viridis(nages,option="D",direction=-1)
                    )#end tibble
}else{
   ageinfo  = tibble( id      = integer(0L)
                    , age_lwr = numeric(0L)
                    , age_upr = numeric(0L)
                    , key     = character(0L)
                    , desc    = character(0L)
                    , labs    = character(0L)
                    , colour  = character(0L)
                    )#end tibble
}#end if (is.na(idxage))

# Set number of age classes
nages = nrow(ageinfo)
}

# List of size classes
if(read_dbh){
idxdbh   = match("fates_levscls",nc_dlist)
if (is.finite(idxdbh)){
   dbhs     = nc_conn$dim[[idxdbh]]$vals
   ndbhs    = nc_conn$dim[[idxdbh]]$len
   dbhinfo  = tibble( id      = sequence(ndbhs)
                    , dbh_lwr = dbhs
                    , dbh_upr = c(dbh_lwr[-1],dbh_lwr[ndbhs]+2*max(diff(dbh_lwr)))
                    , dbh     = 0.5 * (dbh_lwr + dbh_upr)
                    , key     = sprintf("dbh_%3.3g",dbh_lwr)
                    , desc    = c( paste0("paste(paste(",dbh_lwr[-ndbhs],"<=D*B*H)<",dbh_upr[-ndbhs],"*c*m)")
                                 , paste0("paste( D*B*H >=",dbh_lwr[ndbhs],"*c*m)")
                                 )#end c
                    , labs    = c( paste0("paste(",dbh_lwr[-ndbhs],"-",dbh_upr[-ndbhs],")")
                                 , paste0("paste(",dbh_lwr[ ndbhs],"-infinity)")
                                 )#end dbhlabs
                    , colour   = magma(ndbhs,direction=1)
                    )#end tibble
}else{
   dbhinfo  = tibble( id      = integer(0L)
                    , dbh_lwr = numeric(0L)
                    , dbh_upr = numeric(0L)
                    , key     = character(0L)
                    , desc    = character(0L)
                    , labs    = character(0L)
                    , colour  = character(0L)
                    )#end tibble
}#end if (is.finite(idxdbh))

# Set number of size classes
ndbhs = nrow(dbhinfo)
}

# List of PFT classes (only if not using user-defined classes).
idxpft   = match("fates_levpft",nc_dlist)
if (! is.finite(idxpft)){
   # PFT index not found. Skip PFTs altogether.
   pftinfo = tibble( id               = numeric(0L)
                   , key              = character(0L)
                   , short            = character(0L)
                   , desc             = character(0L)
                   , colour           = character(0L)
                   , stringsAsFactors = FALSE
                   )#end data.table
}else if (! user_pftinfo){
   # Select all PFTs available
   pftids  = nc_conn$dim[[idxpft]]$vals
   npftids = nc_conn$dim[[idxpft]]$len

   # Build tibble with all the PFTs.
   pftinfo = tibble( id               = pftids
                   , key              = sprintf("pft%2.2i" ,pftids)
                   , short            = sprintf("PFT%2.2i" ,pftids)
                   , desc             = sprintf("PFT %2.2i",pftids)
                   , colour           = brewer.pal(n=npftids,name="PuBuGn")
                   , stringsAsFactors = FALSE
                   )#end tibble
}else if (! is_tibble(pftinfo)){
   # Convert user-defined pftinfo to a "tibble" object
   pftinfo  = do.call(what=rbind,args=lapply(X=pftinfo,FUN=as_tibble,stringsAsFactors=FALSE))
}#end if (! is.finite(idxpft))

# Set number of PFTs (active PFTs only) 
npfts = nrow(pftinfo)

# List of fuel classes
if(read_fuel){
idxfuel   = match("fates_levfuel",nc_dlist)
if (! is.finite(idxfuel)){
   # fuel index not found. Skip Fuels altogether.
   fuelinfo = tibble( id               = numeric(0L)
                    , key              = character(0L)
                    , short            = character(0L)
                    , desc             = character(0L)
                    , colour           = character(0L)
                    , stringsAsFactors = FALSE
                    )#end data.table
}else if (! user_fuelinfo){
   # Select all fuels available
   fuelids  = nc_conn$dim[[idxfuel]]$vals
   nfuelids = nc_conn$dim[[idxfuel]]$len

   # Build tibble with all the fuels.
   fuelinfo = tibble( id               = fuelids
                    , key              = sprintf("Fuel %2.2i" ,fuelids)
                    , short            = sprintf("Fuel %2.2i" ,fuelids)
                    , desc             = sprintf("Fuel %2.2i" ,fuelids)
                    , colour           = brewer.pal(n=nfuelids,name="PuBuGn")
                    , stringsAsFactors = FALSE
                   )#end tibble
}else if (! is_tibble(fuelinfo)){
   # Convert user-defined fuelinfo to a "tibble" object
   fuelinfo  = do.call(what=rbind,args=lapply(X=fuelinfo,FUN=as_tibble,stringsAsFactors=FALSE))
}#end if (! is.finite(idxfuel))
nfuels = nrow(fuelinfo)
}

if(read_soi){
# List of soil layer
idxsoi   = match("levgrnd",nc_dlist)
if (is.finite(idxsoi)){
   sois     = nc_conn$dim[[idxsoi]]$vals
   nsois    = nc_conn$dim[[idxsoi]]$len
   soikeys  = sprintf("sl_%3.3f",sois)
   soilabs  = c( paste0("paste(",sois[-nsois],"-",sois[-1],")")
               , paste0("paste(",sois[nsois],"-infinity)")
               )#end dbhlabs
   soilabs  = parse(text=soilabs)
}else{
   sois    = numeric(0L)
   nsois   = 0L
   soikeys = character(0L)
   soilabs = character(0L)
}#end if (is.finite(idxdbh))
}

#---~---
#retrieve age-class variables
#---~---
isage    = grepl(pattern="_AP$",x=nc_vlist) | grepl(pattern="_APFC$",x=nc_vlist) |
           grepl(pattern="_APPF$",x=nc_vlist)
nc_byage = nc_vlist[isage]
nc_pref  = gsub(pattern="_AP$",replacement="",x=nc_byage)
nc_pref  = gsub(pattern="_APFC$",replacement="",x=nc_pref)
nc_pref  = gsub(pattern="_APPF$",replacement="",x=nc_pref)
nc_pref  = tolower(nc_pref)
nc_keep  = nc_pref %in% fatesvar_ap$vnam & (! duplicated(nc_pref))
no_byage = nc_byage[! nc_keep]
nc_byage = nc_byage[  nc_keep]
nbyage   = length(nc_byage)


# by size class and whether total LAI can be calculated given canopy
# and under-story LAI

is_size   = grepl(pattern="_SZ$",x=nc_vlist) | grepl(pattern="_SZPF$",x=nc_vlist)
nc_bydbh  = nc_vlist[is_size]
nc_pref   = gsub(pattern="_SZ$",replacement="",x=nc_bydbh)
nc_pref   = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
nc_pref   = tolower(nc_pref)
nc_keep   = (nc_pref %in% fatesvar_szpf$vnam) & (! duplicated(nc_pref))
no_bydbh  = nc_bydbh[! nc_keep]
nc_bydbh  = unique(nc_bydbh[  nc_keep])

vardbh_last = rep(x=FALSE,times=nszpf_read)
for (v in which(fatesvar_szpf$is_upc)){
   nc_vnow      = toupper(fatesvar_szpf$vnam[v])
   nc_vnow_scls = paste0(nc_vnow,"_SZ")
   nc_vund_scls = paste0(nc_vnow,"_USTORY_SZ")
   nc_vcan_scls = paste0(nc_vnow,"_CANOPY_SZ")
   nc_vnow_scpf = paste0(nc_vnow,"_SZPF")
   nc_vund_scpf = paste0(nc_vnow,"_USTORY_SZPF")
   nc_vcan_scpf = paste0(nc_vnow,"_CANOPY_SZPF")

   # Check whether this variable can be derived from understorey+canopy (and needs to be).
   if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) ){
      nc_bydbh       = unique(c(nc_bydbh,nc_vnow_scls))
      vardbh_last[v] = TRUE
   }else if ( all(c(nc_vund_scpf,nc_vcan_scpf) %in% nc_bydbh ) && (! nc_vnow_scpf %in% nc_bydbh) ){
      nc_bydbh       = unique(c(nc_bydbh,nc_vnow_scpf))
      vardbh_last[v] = TRUE
   }#end if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) )
}#end for (v in which(fatesvar$is_upc))


nbydbh    = length(nc_bydbh)
n_vardbh_last = sum(vardbh_last)

#by pft
is_pft    = grepl(pattern="_SZPF$",x=nc_vlist) | grepl(pattern="_PF$",x=nc_vlist) 
nc_bypft  = nc_vlist[is_pft]
nc_pref   = gsub(pattern="_SZPF$",replacement="",x=nc_bypft )
nc_pref   = gsub(pattern="_PF$",replacement="", x=nc_pref  )
nc_pref   = tolower(nc_pref)
nc_keep   = nc_pref %in% fatesvar_new$vnam & (! duplicated(nc_pref))
no_bypft  = nc_bypft[! nc_keep]
nc_bypft  = unique(nc_bypft[nc_keep])

varpft_last = rep(x=FALSE,times=nfatesvar_new)
for (v in which(fatesvar_new$is_upc)){
   nc_vnow      = toupper(fatesvar_new$vnam[v])
   nc_vnow_scpf = paste0(nc_vnow,"_SZPF")
   nc_vund_scpf = paste0(nc_vnow,"_USTORY_SZPF")
   nc_vcan_scpf = paste0(nc_vnow,"_CANOPY_SZPF")

   # Check whether this variable can be derived from understorey+canopy (and needs to be).
   if ( all( c(nc_vund_scpf,nc_vcan_scpf) %in% nc_bypft ) && (! nc_vnow_scpf %in% nc_bypft) ){
      nc_bypft       = unique(c(nc_bypft,nc_vnow_scpf))
      varpft_last[v] = TRUE
   }#end if ( all(c(nc_vund_scls,nc_vcan_scls) %in% nc_bydbh ) && (! nc_vnow_scls %in% nc_bydbh) )
}#end for (v in which(fatesvar$is_upc))

nbypft    = length(nc_bypft)
n_varpft_last = sum(varpft_last)

# by fuel class
nc_byfuel = nc_vlist[grepl(pattern="_FC$",x=nc_vlist)]
nc_pref   = tolower(gsub(pattern="_FC$",replacement="",x=nc_byfuel))
nc_keep   = nc_pref %in% fatesvar_new$vnam & (! duplicated(nc_pref))
no_byfuel = nc_byfuel[! nc_keep]
nc_byfuel = nc_byfuel[  nc_keep]
nbyfuel   = length(nc_byfuel)

#---~---
#    Retrieve all "drought deciduous phenology variables
#---~---
nc_pref   = tolower(x=nc_vlist)
nc_keep   = nc_pref %in% dphenvar$vorig
no_dphen  = nc_vlist[! nc_keep]
nc_dphen  = nc_vlist[  nc_keep]

# Tally the total number of drought phenology variables.
ndphen = length(nc_dphen)


# all 1D and 2D variables variables
nc_pref   = tolower(x=nc_vlist)
nc1d_keep = nc_pref %in% hlm1dvar$vnam
no_hlm1d  = nc_vlist[! nc1d_keep]
nc_hlm1d  = nc_vlist[  nc1d_keep]
nc2d_keep = nc_pref %in% hlm2dsoi$vnam
no_hlm2d  = nc_vlist[! nc2d_keep]
nc_hlm2d  = nc_vlist[  nc2d_keep]

# Check whether to append "evapotranspiration"
if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc_hlm1d) ) && (! "QEVTR" %in% nc_hlm1d) ){
   nc_hlm1d = unique(c(nc_hlm1d,"QEVTR"))
   etr_last = TRUE
}else{
   etr_last = FALSE
}#end if (  ( all(c("QSOIL","QVEGT","QVEGE") %in% nc_hlm1d) ) && (! "QEVTR" %in% nc_hlm1d) )

# Check whether to append ecosystem respiration
if (  ( all(c("FATES_AUTORESP","FATES_HET_RESP") %in% nc_hlm1d) ) && (! "ER" %in% nc_hlm1d) ){
   nc_hlm1d = unique(c(nc_hlm1d,"ER"))
   er_last  = TRUE
}else{
   er_last  = FALSE
}#end if (  ( all(c("AR","HR") %in% nc_hlm1d) ) && (! "ER" %in% nc_hlm1d) )

# Find number of host land model variables
nhlm1d    = length(nc_hlm1d)
nhlm2d    = length(nc_hlm2d)

#---~---
#Initialize variable matrix
#---~---
if(read_age){
#by age class
byage = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  
  for(a in sequence(nbyage)){
  nc_nvnow                   = nc_byage[a]
  nc_pref                    = gsub(pattern="_AP$",replacement="",x=nc_nvnow)
  nc_pref                    = gsub(pattern="_APPF$",replacement="",x=nc_pref)
  nc_pref                    = gsub(pattern="_APFC$",replacement="",x=nc_pref)
  nc_pref                    = tolower(nc_pref)
  f                          = match(nc_pref,fatesvar_ap$vnam)
  f_vnam                     = fatesvar_ap$vnam[f]
  byage[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_each*ninst,ncol=nages,dimnames=list(NULL,ageinfo$key)) 
  } #end of a loop
} #end of c(in sequence(ncase))
}

if(read_dbh){
# by size class
bydbh = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for(d in sequence(nbydbh)){
  nc_nvnow        = nc_bydbh[d]
  nc_pref         = gsub(pattern="_SZ$",replacement="",x=nc_nvnow)
  nc_pref         = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
  nc_pref         = tolower(nc_pref)
  f               = match(nc_pref,fatesvar_szpf$vnam)
  f_vnam          = fatesvar_szpf$vnam[f]
  bydbh[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_each*ninst,ncol=ndbhs,dimnames=list(NULL,dbhinfo$key))
   } # end of d
}
}

# by pft

bypft = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for(p in sequence(nbypft)){
  nc_nvnow        = nc_bypft[p]
  nc_pref         = gsub(pattern="_PF$",replacement="",x=nc_nvnow  )
  nc_pref         = gsub(pattern="_SZPF$",replacement="",x=nc_pref )
  nc_pref         = tolower(nc_pref)
  f               = match(nc_pref,fatesvar_new$vnam)
  f_vnam          = fatesvar_new$vnam[f]
  bypft[[case]][[f_vnam]] = matrix(data=NA_real_,nrow =ntstamp_each*ninst,ncol=npfts,
                           dimnames=list(NULL,pftinfo$key)) 
  } # end of p
}


# Initialize list of variables by fuel class
if(read_fuel){
byfuel = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  
  for(f in sequence(nbyfuel)){
  nc_nvnow                   = nc_byfuel[f]
  nc_pref                    = tolower(gsub(pattern="_FC$",replacement="",x=nc_nvnow))
  f                          = match(nc_pref,fatesvar_new$vnam)
  f_vnam                     = fatesvar_new$vnam[f]
  byfuel[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_each*ninst,ncol=nfuels,dimnames=list(NULL,fuelinfo$key)) 
  } #end of a loop
} #end of c(in sequence(ncase))
}

# Initialise list of variables by soil layer

if(read_phen){
dphen = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  for (p in sequence(ndphen)){
   nc_nvnow        = nc_dphen[p]
   nc_pref         = tolower(nc_nvnow)
   f               = match(nc_pref,dphenvar$vorig)
   f_vnam          = dphenvar$vnam[f]
   dphen[[case]][[f_vnam]] = matrix(data=NA_real_,nrow=ntstamp_each,
                                    ncol=npfts,dimnames=list(NULL,pftinfo$key))
  }#end for (p in sequence(ndphen))
}
}


# Initialize 1D variables available at the HLM
hlm1d = as_tibble( matrix( data     = NA_real_
                         , nrow     = ntstamp_inst*ncase
                         , ncol     = nhlm1d+2
                         , dimnames = list(NULL,tolower(c(nc_hlm1d,"CASE","ENS_LABEL")))
                         )
                 )

# Initialize soil water status related variables available at the HLM for each soil layer into hlm2dvar
if(read_hlm2d){
hlm2d = list()
for(c in sequence(ncase)){
  case     = as.character(case_names[c])
  
  for(m in sequence(nhlm2dsoi)){
  nc_nvnow        = nc_hlm2d[m]
  nc_pref         = tolower(nc_nvnow)
  h               = match(nc_pref,hlm2dsoi$vnam)
  h_vnam          = hlm2dsoi$vnam[h]
  hlm2d[[case]][[h_vnam]] = matrix(data=NA_real_,nrow=ntstamp_inst,ncol=nsois,dimnames=list(NULL,soikeys))
  }
}
}

# Load soil layers
# do a loop to avoid hard-coding here 

if(read_soi){
cat0("   - Load soil information")
slayer = tibble()

for (c in sequence(ncase)) {
case     = as.character(case_names[c])
slayer_now = tibble( zsoi   = c(unlist(ncvar_get(nc=nc_conn, varid='ZSOI',collapse_degen=TRUE))) 
                   , dzsoi  = c(unlist(ncvar_get(nc=nc_conn,varid='DZSOI',collapse_degen=TRUE ))) 
                   , bsw    = c(unlist(ncvar_get(nc=nc_conn,varid='BSW',collapse_degen=TRUE )))  
                   , hksat  = c(unlist(ncvar_get(nc=nc_conn,varid='HKSAT',collapse_degen=TRUE ))) 
                   , sucsat = c(unlist(ncvar_get(nc=nc_conn,varid='SUCSAT',collapse_degen=TRUE )))  
                   , watsat = c(unlist(ncvar_get(nc=nc_conn,varid='WATSAT',collapse_degen=TRUE )))  
                   , case   = c(rep_along(zsoi, case))
               )#end tibble
   slayer = rbind(slayer, slayer_now)
   rm(slayer_now)
}
}

# Load indices
scls   = ncvar_get(nc=nc_conn,varid='fates_scmap_levscpf')
pft    = ncvar_get(nc=nc_conn,varid='fates_pftmap_levscpf')
fuel   = ncvar_get(nc=nc_conn,varid='fates_levfuel')
index_scpf = tibble( scls = scls[,1]
                   , pft = pft[,1]
                   )#end data.table
rm(scls,pft)
# Close connection
dummy   = nc_close(nc_conn) 

```




```{r,label="retrieve-values",message=FALSE,results='hide'}
if ("nc_conn" %in% ls()){dummy = nc_close(nc_conn); rm(nc_conn)}

#for(n in sequence(ntstamp_each)){

 # w_month      = month(each_tstamp[w])  
 # w_year       = year(each_tstamp[w])
  
  # Find conversion factors for monthly variables.
  
   cmon.day = days_in_month(each_tstamp)
   cmon.hr  = day.hr  * cmon.day
   cmon.min = day.min * cmon.day
   cmon.sec = day.sec * cmon.day
   cmon.sec = rep(cmon.sec,time=ninst)

  
# Open NetCDF connection and retrieve variable names

for(c in sequence(ncase)){
  case     = case_names[c]
  nc_conn  = nc_open(case_files[c])
  nc_nvars = nc_conn$nvars
  nc_vlist = rep(NA_character_,times=nc_nvars)
  for (v in sequence(nc_nvars)) nc_vlist[v] = nc_conn$var[[v]]$name

if(read_age){
# Read variables by age, and assign current values to the matrix.  
  for (a in sequence(nbyage)){
    nc_nvnow            = nc_byage[a]
    nc_pref             = gsub(pattern="_AP$",replacement="",x=nc_nvnow)
    nc_pref             = gsub(pattern="_APFC$",replacement="",x=nc_pref)
    nc_pref             = gsub(pattern="_APPF$",replacement="",x=nc_pref)
    nc_pref             = tolower(nc_pref)
    f                   = match(nc_pref,fatesvar_ap$vnam)
    f_vnam              = fatesvar_ap$vnam[f]
    f_add0              = eval(parse(text=fatesvar_ap$add0[f]))
    f_mult              = eval(parse(text=fatesvar_ap$mult[f]))
    nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
    ncm_dat             = nc_dat[,s_t:e_t,] 
    if(nc_pref == "fates_fuel_amount"){
       
       ncm_dat        = aperm(ncm_dat,c(2,3,1))
       dim(ncm_dat)   = c(ninst*ntstamp_each,nages*nfuels)
       f1             = ncm_dat[,1:7]
       f2             = ncm_dat[,8:14]
       f3             = ncm_dat[,15:21]
       f5             = ncm_dat[,29:35]
       f6             = ncm_dat[,36:42]
       ncm_dat        = f1 + f2 + f3 + f5 + f6
       nc_dat         = f_add0 + f_mult * ncm_dat
    }else if(nc_pref == "fates_vegc"){
    ncm_dat             = aperm(ncm_dat,c(2,3,1))
    dim(ncm_dat)        = c(ninst*ntstamp_each,nages*npfts)
    p1                  = ncm_dat[,1:7]
    p2                  = ncm_dat[,8:14]
    p3                  = ncm_dat[,15:21]
    p4                  = ncm_dat[,22:28]
    ncm_dat             = p1 + p2 + p3 + p4
    nc_dat              = f_add0 + f_mult * ncm_dat
    }else{
    ncm_dat             = aperm(ncm_dat,c(2,3,1))
    dim(ncm_dat)        = c(ninst*ntstamp_each,nages)
    nc_dat              = f_add0 + f_mult * ncm_dat
    }
    byage[[case]][[f_vnam]][,]   = nc_dat 
   
}#end for (a in sequence(nbyage))  
}
#---~--- 
   #   Read variables by size, and assign current values to the matrix.  In case LAI
   # is in the list and it is the last variable, we calculate it from canopy and under
   # story, after loading all variables
   #---~---

if(read_dbh){
for (d in sequence(nbydbh-n_vardbh_last)){
      nc_nvnow            = nc_bydbh[d]
      is_szpf             = grepl(pattern="_SZPF$",x=nc_nvnow)
      nc_pref             = gsub(pattern="_SZPF$",replacement="",x=nc_nvnow)
      nc_pref             = gsub(pattern="_SZ$",replacement="",x=nc_pref )
      nc_pref             = tolower(nc_pref)
      f                   = match(nc_pref,fatesvar_szpf$vnam)
      f_vnam              = fatesvar_szpf$vnam[f]
      f_add0              = eval(parse(text=fatesvar_szpf$add0[f]))
      f_mult              = eval(parse(text=fatesvar_szpf$mult[f]))
      f_aggr              = match.fun(fatesvar_szpf$aggr[f])
      nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow)
      ncm_dat             = nc_dat[,s_t:e_t,]
      ncm_dat             = aperm(nc_dat,c(2,3,1))
      dim(ncm_dat)        = c(ninst*ntstamp_each,ndbhs*npfts)
      p1                  = ncm_dat[,1:13]
      p2                  = ncm_dat[,14:26]
      p3                  = ncm_dat[,27:39]
      p4                  = ncm_dat[,40:52]
      ncm_dat             = p1 + p2 + p3 + p4 
      nc_dat              = f_add0 + f_mult * ncm_dat 
      bydbh[[case]][[f_vnam]][] = nc_dat
   }#end for (d in sequence(nbydbh-laidbh_last))

#    Loop through variables to be added last
   for (v in which(vardbh_last)){
      # Retrieve variable and build understory and canopy variables.
      v_vnam = fatesvar_szpf$vnam[v]
      v_vund = paste0(v_vnam,"_ustory")
      v_vcan = paste0(v_vnam,"_canopy")

      # Aggregate data
      bydbh[[case]][[v_vnam]] = bydbh[[case]][[v_vund]] + bydbh[[case]][[v_vcan]]
   }#end for (v in which(vardbh_last))
}  
 

#---~---
# read variables by PFT
#---~---
  for (p in sequence(nbypft-n_varpft_last)){
      # Load variable information
      nc_nvnow = nc_bypft[p]
      is_szpf  = grepl(pattern="_SZPF$",x=nc_nvnow)
      nc_pref  = gsub(pattern="_SZPF$",replacement="",x=nc_nvnow)
      nc_pref  = gsub(pattern="_PF$",replacement="",x=nc_pref )
      nc_pref  = tolower(nc_pref)
      f        = match(nc_pref,fatesvar_new$vnam)
      f_vnam   = fatesvar_new$vnam[f]
      f_add0   = eval(parse(text=fatesvar_new$add0[f]))
      f_mult   = eval(parse(text=fatesvar_new$mult[f]))
      f_aggr   = match.fun(fatesvar_new$aggr[f])
      nc_dat   = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
      ncm_dat   = nc_dat[,s_t:e_t,]
      if(is_szpf){
      # Sum data across size class for each PFT per time step
      ncm_dat      = aperm(ncm_dat,c(2,3,1))
      dim(ncm_dat) = c(ninst*ntstamp_each, ndbhs*npfts)
      p1           = base::rowSums(ncm_dat[,1:13]) # for woody PFT we only sums size class > 2.5cm
      p2           = base::rowSums(ncm_dat[,14:26])
      p3           = base::rowSums(ncm_dat[,27:39])
      p4           = base::rowSums(ncm_dat[,40:52])
      ncm_dat      = rbind(p1,p2,p3,p4)
      ncm_dat      = aperm(ncm_dat,c(2,1))
      nc_dat       = f_add0 + f_mult * ncm_dat
      bypft[[case]][[f_vnam]][] = nc_dat
      }else{
      nc_dat = aperm(nc_dat,c(2,3,1))
      dim(nc_dat) = c(ninst*ntstamp_each, npfts)
      nc_dat = f_add0 + f_mult * nc_dat
      bypft[[case]][[f_vnam]][] = nc_dat
      }
}#end for (p in sequence(nbypft-n_varpft_last))

# Loop through variables to be added last
  for (v in which(varpft_last)){
      # Retrieve variable and build understory and canopy variables.
    v_vnam = fatesvar_new$vnam[v]
    v_vund = paste0(v_vnam,"_ustory")
    v_vcan = paste0(v_vnam,"_canopy")

  # Aggregate data
   bypft[[case]][[v_vnam]] = bypft[[case]][[v_vund]] + bypft[[case]][[v_vcan]]
   }#end for (v in which(varpft_last))
  
if(read_fuel){
   # Read variables by fuel class, and assign current values to the matrix.  
  for (f in sequence(nbyfuel)){
    nc_nvnow            = nc_byfuel[f]
    nc_pref             = tolower(gsub(pattern="_FC$",replacement="",x=nc_nvnow))
    f                   = match(nc_pref,fatesvar_new$vnam)
    f_vnam              = fatesvar_new$vnam[f]
    f_add0              = eval(parse(text=fatesvar_new$add0[f]))
    f_mult              = eval(parse(text=fatesvar_new$mult[f]))
    nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
   #subset time to only include results from 2000 to 2020
    nc_dat              = nc_dat[,s_t:e_t,] 
   #reform the 3d array so when convert to matrix, rows are time by ensemble members, columns are fuel class 
    ncm_dat             = aperm(nc_dat, c(2,3,1)) 
    dim(ncm_dat)        = c(ninst*ntstamp_each,nfuels)
    nc_dat              = f_add0 + f_mult * ncm_dat
    byfuel[[case]][[f_vnam]][,]   = nc_dat
}#end for (f in sequence(nbyfuel))  
}

 # Read drought-deciduous phenology variables (by PFT), and assign current values to the matrix.
   # We use "rev" because the first soil layer is the deepest for the R output.
  if(read_phen){
   for (p in sequence(ndphen)){
      nc_nvnow            = nc_dphen[p]
      nc_pref             = tolower(nc_nvnow)
      f                   = match(nc_pref,dphenvar$vorig)
      f_vnam              = dphenvar$vnam[f]
      f_add0              = eval(parse(text=dphenvar$add0[f]))
      f_mult              = eval(parse(text=dphenvar$mult[f]))
      nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
      dphen[[case]][[f_vnam]][] = f_add0 + f_mult * rev(nc_dat)
      #dphen[[case]][[f_vnam]][,] = f_add0 + f_mult * rev(nc_dat)
   }#end for (a in sequence(nbyage))

}
# Read 1D variables
 
  for (v in sequence(nhlm1d-etr_last-er_last)){
  
  nc_nvnow            = nc_hlm1d[v]
  nc_pref             = tolower(x=nc_nvnow)
  h                   = match(nc_pref,hlm1dvar$vnam)
  h_vnam              = hlm1dvar$vnam[h]
  h_add0              = eval(parse(text=hlm1dvar$add0[h]))
  h_mult              = eval(parse(text=hlm1dvar$mult[h]))
  nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
  nc_dat              = nc_dat[s_t:e_t,] 
  ncm_dat             = matrix(nc_dat, ncol=1)
  
  if (c==1){
         hlm1d[[h_vnam]][1:ntstamp_inst]  = h_add0 + h_mult * nc_dat
         hlm1d$case[1:ntstamp_inst]       = case
         hlm1d$ens_label[1:ntstamp_inst]  = rep(ens_labels,each  = ntstamp_each)
     }else{
        start = (c-1)*ntstamp_inst+1
        end   = c*ntstamp_inst
         hlm1d[[h_vnam]][start:end] = h_add0 + h_mult * nc_dat
         hlm1d$case[start:end]      = case
         hlm1d$ens_label[start:end] = rep(ens_labels,each  = ntstamp_each)
         }
  
}#for (h in sequence(nhlm1d-etr_last-et_last))
  
 # Find total ET.
   if (etr_last){
      if(c==1){
         hlm1d$qevtr[1:ntstamp_inst] = hlm1d$qvege[1:ntstamp_inst] + 
                                       hlm1d$qvegt[1:ntstamp_inst] + 
                                       hlm1d$qsoil[1:ntstamp_inst]
      }else{
         start = (c-1)*ntstamp_inst+1
         end   = c*ntstamp_inst
         hlm1d$qevtr[start:end] = hlm1d$qvege[start:end] + 
                                  hlm1d$qvegt[start:end] + 
                                  hlm1d$qsoil[start:end]
      }
   }#end if (etr_last)

   # Find total ET.
   if (er_last){
      if(c==1){
         hlm1d$er[1:ntstamp_inst] = hlm1d$fates_autoresp[1:ntstamp_inst] + 
                                    hlm1d$fates_het_resp[1:ntstamp_inst]
      }else{
         start = (c-1)*ntstamp_inst+1
         end   = c*ntstamp_inst
         hlm1d$er[start:end] = hlm1d$fates_autoresp[start:end] + 
                               hlm1d$fates_het_resp[start:end] 
               }
   }#end if (er_last)

if(read_hlm2d){
# Read 2D variables
for (m in sequence(nhlm2d)){ 
  nc_nvnow            = nc_hlm2d[m]
  nc_pref             = tolower(x=nc_nvnow)
  h                   = match(nc_pref,hlm2dsoi$vnam)
  h_vnam              = hlm2dsoi$vnam[h]
  h_add0              = eval(parse(text=hlm2dsoi$add0[h]))
  h_mult              = eval(parse(text=hlm2dsoi$mult[h]))
  nc_dat              = ncvar_get(nc=nc_conn,varid=nc_nvnow,collapse_degen = TRUE)
  nc_dat              = nc_dat[,s_t:e_t,] 
  ncm_dat             = aperm(nc_dat, c(2,3,1))
  dim(ncm_dat)        = c(ntstamp_each*ninst, nsois)
  nc_dat              = h_add0 + h_mult * ncm_dat
  hlm2d[[case]][[h_vnam]][,]  = nc_dat
}#end of for (m in sequence(nhlm2d))
}
# Close connection
dummy   = nc_close(nc_conn)

  
} #end of for (c in sequence(ncase))

## add time for hlm1d
hlm1d$time = rep(ens_tstamp,time=ncase)
fwrite(hlm1d,file.path(main_path,paste0(file_desc,"_hlm1d.csv")))

```




```{r,label="scale-age",message=FALSE,results='hide'}
cat0(" + Rescale variables by age class.")

for(c in sequence(ncase)){
  case = case_names[c]
  for (a in sequence(nbyage)){
#--- Match variables
  f        = match(names(byage[[case]])[a],fatesvar_ap$vnam)
  f_vnam   = fatesvar_ap$vnam [f]
  f_desc   = fatesvar_ap$desc [f]
  f_stack  = fatesvar_ap$stack[f]
   #---~---

   #--- Proceed only if the variable is stacked and not the patch area.
  if (f_stack && (! (f_vnam %in% "fates_patcharea"))){
      cat0("   - ",f_desc,".")
      byage[[case]][[f_vnam]] = byage[[case]][[f_vnam]] * byage[[case]]$fates_patcharea
  }#end if (f_stack && (! (f_vnam %in% "patch_area")))
}#end for (a in sequence(nbyage))
  
}



```




```{r,label="melt-data",message=FALSE,results='hide'}
cat0(" + Turn age-dependent matrices into tibble objects.")
age_melt  = NULL

for (c in sequence(ncase)){
  case = case_names[c]
  now_agemelt = NULL
  for (a in sequence(nbyage)){
   #--- Match variables.
  f        = match(names(byage[[case]])[a],fatesvar_ap$vnam)
  f_vnam   = fatesvar_ap$vnam[f]
  f_desc   = fatesvar_ap$desc[f]
  cat0("   - ",f_desc,".")
      
#--- Create molten data table for this variable.  
  now_age           = as_tibble(byage[[case]][[f_vnam]])
  now_age$time      = rep(each_tstamp,time=ninst)
  now_age$case      = rep_along(time,x = case)
  now_age$ens_label = rep(ens_labels,each = ntstamp_each)
  now_melt          = as_tibble(reshape2::melt(data=now_age,id.vars=c("case", "ens_label",    
                              "time"),variable.name="age",value.name=f_vnam))
      #--- Merge data table
  
  if (is.null(now_agemelt)){
    now_agemelt = now_melt
  }else{
        now_agemelt = now_agemelt %>% left_join(now_melt, by=c("case","time","age","ens_label"))
  }#end if (is.null(age_melt))
  
  }#end for (a in sequence(nbyage))
age_melt = rbind(age_melt,now_agemelt)
}



byage = age_melt 
rm(age_melt)
invisible(gc())

cat0(" + Turn size-dependent matrices into data tables.")
dbh_melt = NULL
for (c in sequence(ncase)){
  case        = case_names[c]
  now_dbhmelt = NULL
for (d in sequence(nbydbh)){
  f        = match(names(bydbh[[case]])[d],fatesvar_szpf$vnam)
  f_vnam   = fatesvar_szpf$vnam[f]
  f_desc   = fatesvar_szpf$desc[f]
  cat0("   - ",f_desc,".")
   
  now_dbh           = as_tibble(bydbh[[case]][[f_vnam]])
  now_dbh$time      = rep(each_tstamp, time=ninst)
  now_dbh$case      = rep_along(time, x = case)
  now_dbh$ens_label = rep(ens_labels, each = ntstamp_each)
  now_melt          = as_tibble(reshape2::melt(data=now_dbh,id.vars=c("case","ens_label", "time"),variable.name="dbh",value.name=f_vnam))
  
  if (is.null(now_dbhmelt)){
      now_dbhmelt = now_melt
      }else{
     # now_dbhmelt = merge(x=now_dbhmelt,y=now_melt,by=c("case","time","dbh"),all=TRUE)
      now_dbhmelt = now_dbhmelt %>% left_join(now_melt,by=c("case","ens_label","time","dbh"))
      }#end if (is.null(scls.melt))
}#end for (d in sequence(nbydbh))
dbh_melt = rbind(dbh_melt,now_dbhmelt)
}

bydbh = dbh_melt
rm(dbh_melt)
invisible(gc())

cat0(" + Turn PFT-dependent matrices into data tables.")
pft_melt = NULL
for (c in sequence(ncase)){
  case = case_names[c]
  now_pftmelt = NULL
  for (p in sequence(nbypft)){
  f        = match(names(bypft[[case]])[p],fatesvar_new$vnam)
  f_vnam   = fatesvar_new$vnam[f]
  f_desc   = fatesvar_new$desc[f]
  cat0("   - ",f_desc,".")
  
  now_pft           = as_tibble(bypft[[case]][[f_vnam]])
  now_pft$time      = rep(each_tstamp, time=ninst)
  now_pft$case      = rep_along(time, x=case)
  now_pft$ens_label = rep(ens_labels, each=ntstamp_each)
  now_melt          = as_tibble(reshape2::melt(data=now_pft,id.vars=c("case","ens_label",
                                               "time"),variable.name="pft",value.name=f_vnam))

  if (is.null(now_pftmelt)){
      now_pftmelt = now_melt
      }else{
      #now_pftmelt = as_tibble(merge(x=now_pftmelt,y=now_melt,by=c("case","time", "pft"),all=TRUE))
       now_pftmelt = now_pftmelt %>% left_join(now_melt, by=c("case","ens_label","time","pft"))
      }
}#end for (p in sequence(nbypft))
  pft_melt = rbind(pft_melt,now_pftmelt)
}


bypft = pft_melt
rm(pft_melt)
invisible(gc())

cat0(" + Turn fuel-class matrices into tibble objects.")
fuel_melt  = NULL

for (c in sequence(ncase)){
  case = case_names[c]
  now_fuelmelt = NULL
  for (u in sequence(nbyfuel)){
   #--- Match variables.
  f        = match(names(byfuel[[case]])[u],fatesvar_new$vnam)
  f_vnam   = fatesvar_new$vnam[f]
  f_desc   = fatesvar_new$desc[f]
  cat0("   - ",f_desc,".")
      
#--- Create molten data table for this variable.  
  now_fuel            = as_tibble(byfuel[[case]][[f_vnam]])
  now_fuel$time       = rep(each_tstamp, time=ninst)
  now_fuel$case       = rep_along(time, x = case)
  now_fuel$ens_label  = rep(ens_labels, each=ntstamp_each)
  now_melt            = as_tibble(reshape2::melt(data=now_fuel,id.vars=c("case","ens_label",      
                              "time"),variable.name="fuel",value.name=f_vnam))
      #--- Merge data table
  
  if (is.null(now_fuelmelt)){
    now_fuelmelt = now_melt
  }else{
        #now_fuelmelt = as_tibble(merge(x=now_fuelmelt,y=now_melt,by=c("case","time","fuel"),all=TRUE))
        now_fuelmelt = now_fuelmelt %>% left_join(now_melt, by=c("case","ens_label","time","fuel"))
  }#end if (is.null(age_melt))
  
  }#end for (a in sequence(nbyage))
fuel_melt = rbind(fuel_melt,now_fuelmelt)
}



byfuel = fuel_melt 
rm(fuel_melt)
invisible(gc())

# Turn drought-deciduous phenology variables into data tables
if (! is_tibble(dphen)){
   cat0(" + Turn drought-deciduous phenology matrices into data tables.")
   dph_melt = NULL
   for (c in sequence(ncase)){
     case = case_names[c]
     now_dphmelt = NULL
     for (p in sequence(ndphen)){
      # Match variables.
      f        = match(names(dphen[[case]])[p],dphenvar$vnam)
      f_vnam   = dphenvar$vnam[f]
      f_desc   = dphenvar$desc[f]
      cat0("   - ",f_desc,".")

      # Create molten data table for this variable.  
      now_dph           = as_tibble(dphen[[case]][[p]])
      now_dph$time      = each_tstamp
      now_dph$case      = rep_along(time, x=case)
      now_melt          = as_tibble(melt(data=now_dph,id.vars=c("case","time"),
                                    variable.name="pft",value.name=f_vnam))

      # Merge data table
      if (is.null(now_dphmelt)){
         now_dphmelt = now_melt
      }else{
         now_dphmelt = as_tibble(merge(x=now_dphmelt,y=now_melt,by=c("case","time","pft"),all=TRUE))
      }#end if (is.null(dph_melt))
   }#end for (p in sequence(ndphen))
    dph_melt = rbind(dph_melt,now_dphmelt)
}# end of for (c insequence(ncase))
  
   # Replace dphen with the tibble object
   dphen = dph_melt
   rm(dph_melt)
   invisible(gc())
}#end if (! is_tibble(dphen))

cat0(" + Turn soil layer-dependent matrices into data tables.")
soi_melt = NULL
for (c in sequence(ncase)){
  case = case_names[c]
  now_soimelt = NULL
  for (s in sequence(nhlm2d)){
  h        = match(names(hlm2d[[case]])[s],hlm2dsoi$vnam)
  h_vnam   = hlm2dsoi$vnam[h]
  h_desc   = hlm2dsoi$desc[h]
  cat0("   - ",h_desc,".")
   
  now_soi           = as_tibble(hlm2d[[case]][[s]])
  now_soi$time      = each_tstamp
  now_soi$case      = rep_along(time, x=case)
  now_melt          = as_tibble(reshape2::melt(data=now_soi,id.vars=c("case",              
                                             ,"time"),variable.name="soi",value.name=h_vnam))
  if (is.null(now_soimelt)){
      now_soimelt = now_melt
      }else{
         now_soimelt = merge(x=now_soimelt,y=now_melt,by=c("case","time","soi"),all=TRUE)
      }#end if (is.null(scls.melt))
  }#end for (s in sequence(nsoi))
  soi_melt = rbind(soi_melt,now_soimelt)
  
}


bysoi = soi_melt
rm(soi_melt)
invisible(gc())


cat0(" + Convert classes to integers.")
byage$age = as.integer(byage$age)
bydbh$dbh = as.integer(bydbh$dbh)
bypft$pft = as.integer(bypft$pft)
#dphen$pft = as.integer(dphen$pft)
#bysoi$soi = as.integer(bysoi$soi)
byfuel$fuel = as.integer(byfuel$fuel)

## write out tables so can use later
fwrite(byage,file.path(main_path,paste0(file_desc,"_byage.csv")))
fwrite(bydbh,file.path(main_path,paste0(file_desc,"_bydbh.csv")))
fwrite(bypft,file.path(main_path,paste0(file_desc,"_bypft.csv")))
#fwrite(dphen,file.path(main_path,paste0("Tonzi-2pft-AVBA-280_dphen.csv")))
#fwrite(bysoi,file.path(main_path,paste0("Tonzi-2pft-AVBA-280_bysoi.csv")))
fwrite(byfuel,file.path(main_path,paste0(file_desc,"_byfuel.csv")))

bypft  = fread(file.path(main_path,paste0(file_desc,"_bypft.csv")))
hlm1d  = fread(file.path(main_path,paste0(file_desc,"_hlm1d.csv")))
bydbh  = fread(file.path(main_path,paste0(file_desc,"_bydbh.csv")))
byfuel = fread(file.path(main_path,paste0(file_desc,"_byfuel.csv")))

```





```{r,label="load-site-data",message=FALSE,results='hide'}
cat0(" + Load site data from ",obs_main,".")
tonzi_obs  = fread(flux_path)
tonzi_obs  = tonzi_obs               %>% 
             mutate(fsh         = FSH_can + FSH_und,
                    eflx_lh_tot = EFLX_LH_TOT_can + EFLX_LH_TOT_und,
                    gpp         = GPP_can + GPP_und) %>% 
             dplyr::select(time, fsh, eflx_lh_tot,gpp)

ca_lai   = fread(lai_path)
oak_mas  = fread(mas_path)
oak_can  = fread(can_path)
obs_name = c("ca_lai","oak_mas","oak_can")
obs_desc = c("lai","biomass","cancov")
n_obs    = length(obs_desc)

ton_cord  = data.frame(x = -120.966, y   = 38.4309)
all_obs   = NULL
for(s in sequence(ncase)){
  case_now = case_names[s]
  cord_now = ton_cord
  obs_lst  = NULL
  for(v in sequence(n_obs)){
    obn    = obs_name[v]
    obd    = obs_desc[v]
    ob_now = eval(parse(text = obn))
    nn     = RANN::nn2(ob_now[,1:2],cord_now,k=1)
    ob_now$id = as.vector(nn$nn.idx)
    ob_now = ob_now                                 %>% 
             filter(lon==lon[id] & lat==lat[id])    %>% 
             dplyr::select(-c(lon,lat,id))       
    ob_now$case = case_now
    if (is.null(obs_lst)){
     obs_lst = ob_now
    }else if(obn %in% c("ca_lai","ca_gpp")){
    obs_lst  = merge(x=obs_lst,y=ob_now,by=c("case","year","month"),all=TRUE)
   }else{obs_lst = merge(x=obs_lst,y=ob_now,by="case",all=TRUE)}
    
  }
all_obs = rbind(all_obs,obs_lst)
}

all_obs  = all_obs  %>%  mutate(time = make_datetime(year=year,month=month,day=1L))
all_obs  = tonzi_obs %>% left_join(all_obs, by="time")

```




```{r,label='time-intersect',message=FALSE,results='hide'}

tstampa_com = max(c(min(each_tstamp),min(all_obs$time)))
tstampz_com = min(c(max(each_tstamp),max(all_obs$time)))

tstamp_com  = each_tstamp[(each_tstamp >= tstampa_com) & (each_tstamp <= tstampz_com)]
ntstamp_com = length(tstamp_com)

obs_mmean = all_obs                          %>% 
             dplyr::select(-c(year,time))    %>% 
             group_by(case,month)            %>% 
             summarise_all(mean,na.rm=TRUE)  %>% 
             ungroup()                       %>% 
             filter(!is.na(month))

#upper and lower range of site observations
obs_vars = names(obs_mmean)[-c(1:2)] 
obs_mupr =  all_obs                                                           %>% 
            group_by(case,month)                                              %>%
            dplyr::select(-c(year,time))                                      %>%  
            summarise_all(quantile,probs=qupr_ribbon,names=FALSE, na.rm=TRUE) %>%
            ungroup()                                                         %>%
            filter(!is.na(month))                                             %>% 
            rename_at(vars(obs_vars),function(x) paste0(x,"_qupr"))

obs_mlwr =  all_obs                                                           %>% 
            group_by(case,month)                                              %>%
            dplyr::select(-c(year,time))                                      %>%   
            summarise_all(quantile,probs=qlwr_ribbon,names=FALSE, na.rm=TRUE) %>%
            ungroup()                                                         %>%
            filter(!is.na(month))                                             %>%
            rename_at(vars(obs_vars),function(x) paste0(x,"_qlwr"))

obs_mmean = merge(x=obs_mmean,y=merge(x=obs_mupr,y=obs_mlwr,by=c("case","month"))
                  ,by=c("case","month"))      


#have to rename gpp and agb so names match for making seasonal dynamic plots
obs_vars = names(obs_mmean)
obs_vars = gsub(pattern="gpp",replacement="fates_gpp",x=obs_vars)
obs_vars = gsub(pattern="lai",replacement="elai",x=obs_vars)
obs_vars = gsub(pattern="biomass",replacement="fates_vegc_abovegorund",x=obs_vars)
names(obs_mmean) = obs_vars


```




```{r,label="pft-summary-datasets",fig.height=50, fig.width=40,message=FALSE}
# if agb during peak growing season (5,6,7)>= 4 .3 kgC, grass > 0.05;we mark the ensemble as survived
# later we can calculate the (mean - sd) using site obs as threshold
growing_sea = c("4","5","6","7","8")

mortout_pft    = names(bypft)[grep(pattern="mortality",x=names(bypft))]
mortout_dbh    = names(bydbh)[grep(pattern="mortality",x=names(bydbh))]


find_mort = function(x,n){
  mort_mon = ifelse( test = n == 0, yes = NA, no = pmin(1.,x/n/12.))
  mort_year = ifelse(is.na(mort_mon), NA, 100. * (1. - (1. - mort_mon)^12))
  return(mort_year)
}#end function find_mort


bypft  =     bypft                                              %>% 
             mutate(month = month(time)
                   ,year  = year(time)
                   ,desc  = case_desc[bypft$case]
                   ,case_ens = paste0(case,"_",ens_label))      %>% 
             mutate_at(c(mortout_pft),
                       ~ find_mort(x=.x,n= if_else(.x=="fates_mortality_ustory", 
                                                   .data$fates_nplant_ustory, 
                                                   .data$fates_nplant)))

bydbh      = bydbh                                              %>% 
             mutate(month = month(time),
                    year  = year(time),
                    desc  = case_desc[bydbh$case],
                    case_ens = paste0(case,"_",ens_label))      %>% 
             mutate_at(c(mortout_dbh),
                       ~find_mort(x=.x,n= if_else(.x=="fates_mortality_ustory", 
                                                   .data$fates_nplant_ustory, 
                                                   .data$fates_nplant)))
  
pft_mmean =   bypft                                                 %>% 
              filter(year %in% c(lst_s:lst_e))                      %>% 
              group_by(case,ens_label,case_ens,desc,pft,month)      %>% 
              dplyr::select(-c(year,time))                          %>% 
              summarise_all(mean, na.rm=TRUE)                       %>% 
              ungroup()


pft_amean =   bypft                                                 %>% 
              group_by(case,ens_label,case_ens,desc,pft,year)       %>% 
              dplyr::select(-c(time,month))                         %>% 
              summarise_all(mean,na.rm=TRUE)                        %>%
              ungroup()


pft_strucmean =    bypft                                            %>% 
              group_by(case,ens_label,case_ens,desc,pft)            %>% 
              filter(year %in% c(lst_s:lst_e))                      %>% 
              dplyr::select(-c(time,month,year))                    %>% 
              summarise_all(mean,na.rm=TRUE)                        %>% 
              ungroup()

pft_mortmean =    bypft                                             %>%
              select(all_of(c("case","ens_label","case_ens",
                              "desc","pft",mortout_pft)))           %>% 
              group_by(case,ens_label,case_ens,desc,pft)            %>% 
              summarise_all(mean,na.rm=TRUE)                        %>% 
              ungroup()
pft_mortmean = pft_mortmean               %>% 
               melt(id.vars = c("case","ens_label","case_ens","desc","pft"),
                    variable.name = "mortality")

  
```





```{r,label="ensemble-filtering-fireon"}

ens_filter =   pft_strucmean                                %>% 
               group_by(case,ens_label,case_ens)            %>% 
               mutate(survi_lab = (fates_vegc_aboveground[pft==1] >1 &
                                   fates_vegc_aboveground[pft==2] >1 &
                                   fates_vegc_aboveground[pft==3] >1 &
                                   fates_vegc_aboveground[pft==4] >0)) %>% 
               filter(survi_lab) %>% ungroup()

c1_filter  = ens_filter                                      %>% 
             filter(case==1)                                 %>% 
             group_by(case,ens_label,case_ens)               %>% 
             mutate(survi_lab = fates_basalarea[pft==1] > fates_basalarea[pft==3]) %>% 
             ungroup()                                       %>% 
             filter(survi_lab)

c2_filter  = ens_filter                                      %>% 
             filter(case==2)                                 %>% 
             group_by(case,ens_label,case_ens)               %>% 
             mutate(survi_lab = fates_basalarea[pft==1] < fates_basalarea[pft==3]) %>% 
             ungroup()                                       %>% 
             filter(survi_lab)

survi_lab = intersect(unique(c1_filter$ens_label), unique(c2_filter$ens_label))
survi_ens = unique(ens_filter$case_ens[ens_filter$ens_label %in% survi_lab])
ens_filter = ens_filter %>% filter(case_ens %in% survi_ens)
nsurvi    = length(survi_lab)

```





```{r,label="ensemble-filtering-fireoff"}

## PFT coexistence 
#ens_filter =   pft_strucmean                                %>% 
#               group_by(case,ens_label,case_ens)            %>% 
#               mutate(survi_lab = (fates_vegc_aboveground[pft==1] >0.5 &
#                                   fates_vegc_aboveground[pft==2] >0.5 &
#                                   fates_vegc_aboveground[pft==3] >0.5 &
#                                   fates_vegc_aboveground[pft==4] >0)) %>% 
#               filter(survi_lab) %>% ungroup()

# relative composition

ens_filter = pft_strucmean                                                           %>% 
             group_by(case,ens_label,case_ens)                                       %>% 
             mutate(sum_ba = sum(fates_basalarea),
                    oak_bapct = fates_basalarea[pft==4] / sum_ba)                    %>% 
             mutate(survi_lab = ((fates_basalarea[pft==1] < fates_basalarea[pft==2] |
                                 fates_basalarea[pft==1] < fates_basalarea[pft==3]) &
                                 oak_bapct < 0.05))          %>% 
             ungroup()                                       %>% 
             filter(survi_lab)

## site characteristics
ens_filter = ens_filter                                     %>% 
             group_by(case,ens_label,case_ens)               %>% 
             mutate(sum_agb   = sum(fates_vegc_aboveground),
                    sum_stem  = sum(fates_nplant),
                    sum_ste_uns = sum(fates_nplant_ustory),
                    fir_stempct = fates_nplant[pft==3]/sum_stem,
                    fir_uns_stempct = fates_nplant_ustory[pft==3]/sum_ste_uns)%>% 
             mutate(survi_lab = (#sum_ba >= 10.5 &
                                 #sum_ba <= 63.3 &
                                 #sum_agb>= 2.   &
                                 #sum_agb<= 16   &
                                 #sum_stem>268    &
                                 fir_stempct>0.25 &
                                 fir_uns_stempct > 0.5))     %>% 
             ungroup()                                       %>% 
             filter(survi_lab)
## meet all criteria at both sites
survi_lab = intersect(unique(ens_filter$ens_label[ens_filter$case==1]), 
                      unique(ens_filter$ens_label[ens_filter$case==2]))

survi_manual = c("node_001-task_007","node_001-task_021",
                 "node_002-task_023","node_002-task_034",
                 "node_002-task_036")
survi_lab    =c(survi_lab,survi_manual)
survi_ens = unique(pft_strucmean$case_ens[pft_strucmean$ens_label %in% survi_lab])
ens_filter = pft_strucmean %>% filter(case_ens %in% survi_ens)
nsurvi    = length(survi_lab)

```




```{r,label="plot-pft-time-series",message=FALSE,results="hide"}

# Title for legend
pft_legend = "Plant functional types"


pft_loop = which(fatesvar_new$vnam %in% names(bypft))
gg_pft   = list()
for(c in sequence(ncase)){
  case_now = case_names[c]
  desc_now = case_desc[c]
  for (f in pft_loop){
  # Match variables.
  f_vnam   = fatesvar_new$vnam [f]
  f_desc   = fatesvar_new$desc [f]
  f_unit   = fatesvar_new$unit [f]
  f_stack  = fatesvar_new$stack[f]
  #f_stack  = FALSE
  cat0("   - ",f_desc,".")

  # Set plotting characteristics.
  f_bypft     = bypft %>% filter(case==case_now & case_ens %in% survi_ens)
  f_bypft$pft = factor(f_bypft$pft,levels=sequence(npfts))
  f_colpfts   = pftinfo$colour
  f_pftlabs   = pftinfo$short

  # Initialise plot (decide whether to plot lines or stacks).
  if (f_stack){
     gg_now = ggplot(data=f_bypft,aes_string(x="time",y=f_vnam,group="pft",fill="pft"))
     gg_now = gg_now + facet_wrap(.~case_ens)
     gg_now = gg_now + scale_fill_manual(name=pft_legend,labels=f_pftlabs,values=f_colpfts)
     gg_now = gg_now + geom_area(position=position_stack(reverse = TRUE),show.legend = TRUE)
  }else{
     gg_now = ggplot(data=f_bypft,aes_string(x="time",y=f_vnam,group="pft",colour="pft"))
     #gg_now = gg_now + facet_wrap(.~case_ens)
     gg_now = gg_now + scale_colour_manual(name=pft_legend,labels=f_pftlabs,values=f_colpfts)
     gg_now = gg_now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f_stack)
  gg_now = gg_now + labs(title=desc_now)
  gg_now = gg_now + scale_x_datetime(date_labels=gg_tfmt)
  gg_now = gg_now + xlab("Simulation time")
  gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , plot.title        = element_text( size = gg_ptsz)
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme

  # Save plot in every format requested.
  for (d in sequence(ndevice)){
    f_output = paste0(f_vnam,"-tspft-",case_now,".",gg_device[d])
    dummy = ggsave( filename = f_output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tspft_path
                  , width    = gg_widthn*0.75
                  , height   = gg_height*0.65
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))

  # Write plot settings to the list.
  gg_pft[[f_vnam]] = gg_now
}#end for (f in pft_loop)
  
}




# If sought, plot images on screen
if (gg_screen) gg_pft


```





```{r,label="plot-pft-seasonal-dynamic",message=FALSE,results='hide'}

cat0(" + Plot seasonal dynamics of each PFT for survived ensembles.")

plot_season = c("fates_gpp","fates_vegc_aboveground","fates_lai","fates_mortality_fire",
              "fates_npp", "fates_mortality_rxcambial", "fates_mortality_cambialburn",
              "fates_mortality_crownscorch","fates_mortality_rxcrown","fates_mortality_rxfire",
              "fates_mortality_ustory","fates_basalarea","fates_seed_alloc")
season_loop = which(fatesvar_new$vnam %in% plot_season)

gg_season   = list()
for (c in sequence(ncase)){
  case_now = case_names[c]
  desc_now = case_desc[c]
  plot_df  = pft_mmean %>% filter(case==case_now & case_ens %in% survi_ens)
  plot_df$pft = factor(plot_df$pft,levels=sequence(npfts))
  for (s in season_loop){
  # Match variables.
  s_vnam   = fatesvar_new$vnam [s]
  s_desc   = fatesvar_new$desc [s]
  s_unit   = fatesvar_new$unit [s]
  s_stack  = fatesvar_new$stack[s]
  #f_stack  = FALSE
  cat0("   - ",s_desc,".")

  # Set plotting characteristics.
 
  s_colpfts   = pftinfo$colour
  s_pftlabs   = pftinfo$short
  names(s_pftlabs) = c("1","2","3","4")

  # Initialise plot.
 
  gg_now = ggplot(data=plot_df,aes_string(x="month",y=s_vnam,colour="pft"))
  gg_now = gg_now + facet_wrap(.~ens_label)
  gg_now = gg_now + scale_colour_manual(values=s_colpfts,labels=s_pftlabs)
  gg_now = gg_now + geom_line(linewidth=2,show.legend = TRUE,alpha=0.5)

  gg_now = gg_now + labs(title=desc_now)
  gg_now = gg_now + scale_x_continuous( breaks = sequence(12))
  gg_now = gg_now + xlab("")
  gg_now = gg_now + ylab(desc.unit(desc=s_desc,unit=untab[[s_unit]],twolines=TRUE))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "right"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , plot.title        = element_text( size = gg_ptsz)
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme

  # Save plot in every format requested.
  for (d in sequence(ndevice)){
    f_output = paste0(case_now,s_vnam,"-tspft-",".",gg_device[d])
    dummy = ggsave( filename = f_output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = secy_path
                  , width    = gg_widthn*0.75
                  , height   = gg_height*0.65
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))

  # Write plot settings to the list.
  gg_season[[s_vnam]] = gg_now
}#end for (f in pft_loop)
  
}



# If sought, plot images on screen
#if (gg_screen) gg_season






```



```{r,label="model-site-seasonal-mean",results='hide'}
hlm_var   = c("fates_gpp","fates_vegc_aboveground","fates_fuel_amount",
              "eflx_lh_tot","elai","fsh","fates_burnfrac","fates_fire_intensity",
              "fates_rxfire_burnfrac", "fates_rxfire_intensity_burnfrac",
              "fates_fire_intensity_burnfrac", "fates_rx_burn_window",
              "rain", "q2m", "u10", "tsa","snow")

hlm_scy   = hlm1d                                                                  %>% 
            dplyr::select(all_of(c("case","ens_label","time",
                                   hlm_var)))                                      %>% 
            mutate(#fates_fire_intensity_burnfrac = ifelse(fates_burnfrac == 0, 0,
                  #        fates_fire_intensity_burnfrac / fates_burnfrac),
                   #fates_rxfire_intensity_burnfrac = ifelse(fates_rxfire_burnfrac == 0, 0,
                   #       fates_rxfire_intensity_burnfrac / fates_rxfire_burnfrac),
                   rain = rain + snow,
                   case_ens = paste0(case,"_",ens_label))

#var_both = intersect(names(obs_mmean),names(hlm_scy))
#var_both = var_both[!var_both %in% c("case","time")]


mmean_com = hlm_scy                                                   %>%
            filter(case_ens %in% survi_ens)                           %>% 
            mutate(month = month(time),
                   year  = year(time))                                %>%
            filter (year %in% c(lst_s:lst_e))                         %>% 
            group_by(month,case,ens_label,case_ens)                   %>%
            dplyr::select(-c(time,year))                              %>%
            summarise_all(mean, na.rm=TRUE)                           %>%
            ungroup()    

mmean_com = mmean_com %>% arrange(case,ens_label,case_ens,month)

mean_com = hlm_scy                                                    %>%
            mutate(month = month(time),
                   year  = year(time))                                %>%
            filter (year %in% c(lst_s:lst_e) &
                    case_ens %in% survi_ens)                          %>% 
            group_by(case,ens_label,case_ens)                         %>%
            dplyr::select(-c(time,month,year))                        %>%
            summarise_all(mean, na.rm=TRUE)                           %>%
            ungroup()    

mean_com = mean_com %>% arrange(case,ens_label,case_ens)


```




```{r,label="model-observation-monthly-comparison"}
cat0(" + Plot monthly mean of HLM 1D variables.")

#leg_colours = c(site_colour,ens_colour)
#leg_labels  = c("Site",case_desc)

emean_loop = which(hlm1dvar$vnam %in% hlm_var)

gg_emean   = list()
for (h in emean_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_vqlwr  = paste0(h_vnam,"_qlwr")
   h_vqupr  = paste0(h_vnam,"_qupr")
   h_legend = v == 1
   cat0("   - ",h_desc,".")

   mmean_now        = mmean_com
   #obs              = obs_mmean 
   mmean_now$case = factor(mmean_now$case,levels=unique(mmean_now$case))

   
   gg_now = ggplot(data=mmean_now)
   gg_now = gg_now + facet_wrap(.~case,labeller=as_labeller(case_desc))
   gg_now = gg_now + geom_line(aes_string(x="month",y=h_vnam,group="ens_label",color="ens_label")
                              ,linewidth=1.5,alpha=0.5,show.legend=TRUE)
   #gg_now = gg_now + geom_line(data=obs
    #                          ,aes_string(x="month",y=h_vnam,color="'Site'")
    #                          ,show.legend = TRUE,linewidth=1.5)
   gg_now = gg_now + scale_colour_manual(name="",
                                         values=c(top12))
                                                 #,"Site" = site_colour))
   #gg_now = gg_now + geom_ribbon( data = obs
  #                              , aes_string(x = "month",ymin=h_vqlwr,ymax=h_vqupr)
  #                              , alpha       = alpha_ribbon
  #                              , inherit.aes = FALSE
  #                              , colour      = "transparent"
  #                              )
   gg_now = gg_now + labs(title=h_desc)
   gg_now = gg_now + scale_x_continuous( breaks = sequence(12))#end scale_x_continuous
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,".",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_widthn
                    , height   = gg_height
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  gg_emean[[h_vnam]] = gg_now      

}


# If sought, plot images on screen
if (gg_screen) gg_emean

```



```{r,label="climate-structure-relation"}

mean_loop = which(hlm1dvar$vnam %in% hlm_var)

gg_emean   = list()
for (h in mean_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_legend = v == 1
   cat0("   - ",h_desc,".")

   mmean_now      = mean_com 
  # mmean_now$case = factor(mmean_now$case,levels=unique(mmean_now$case))
   
   gg_now = ggplot(data=mmean_now)
  # gg_now = gg_now + facet_wrap(.~case,labeller=as_labeller(case_desc))
   gg_now = gg_now + geom_line(aes_string(x="rain",y=h_vnam,color="param")
                              ,linewidth=1.5,alpha=0.5,show.legend=TRUE)
   gg_now = gg_now + scale_colour_manual(name="",
                                         values=ens_colour)
   gg_now = gg_now + labs(title=h_desc)
   gg_now = gg_now + xlab(expression("Precipitation ("~mm~month^-1~")"))
   #gg_now = gg_now + xlab(expression("Two-meter temperature ("~degree*C*~")"))
   gg_now = gg_now + ylab(desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,"rain.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_width*0.4
                    , height   = gg_height*0.5
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  gg_emean[[h_vnam]] = gg_now      

}

```




```{r,label="plot-PFT-composition"}
comp_var = c("fates_crownarea","fates_basalarea","fates_npp",
             "fates_nplant","fates_vegc_aboveground","fates_seed_alloc",
             "fates_nplant_canopy","fates_nplant_ustory")
comp_loop = which(fatesvar_new$vnam %in% comp_var)

for (c in comp_loop){
  f_vnam   = fatesvar_new$vnam[c]
  f_desc   = fatesvar_new$desc[c]
  f_unit   = fatesvar_new$unit[c]
  f_legend = v == 1
  s_colpfts   = pftinfo$colour
  s_pftlabs   = pftinfo$short
  names(s_pftlabs) = c("1","2","3","4")
  
  cat0("   - ",f_desc,".")
  df_now = pft_strucmean %>% filter(case_ens %in% survi_ens)
  #df_now$node = substr(df_now$ens_label,8,8)
  #df_now$task = substr(df_now$ens_label,16,17)
  #tk = c("01","02","03","04","05","06","07","08")
  #tk = c("09","10","11","12","13","14","15","16")
  #tk = c("17","18","19","20","21","22","23","24")
  #tk = c("25","26","27","28","29","30","31","32")
  #tk  =c("33","34","35","36")
  #df_now      = df_now %>% filter(node==2 & task%in% tk)
  df_now$pft   = factor(df_now$pft,levels=sequence(npfts))
  df_now$case  = factor(df_now$case,levels=sequence(ncase))
  gg_now = ggplot(df_now,aes_string(x = "ens_label", y = f_vnam, fill="pft"))
  gg_now = gg_now + facet_wrap(.~case,labeller = labeller(case=case_desc))
  gg_now = gg_now + geom_bar(stat = "identity")
  gg_now = gg_now + scale_fill_manual(values = s_colpfts,labels= s_pftlabs )
  gg_now = gg_now + xlab("")
  gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                            , angle=90
                                                            , hjust=1
                                                            , margin = unit(rep(0.35,times=4),"cm"))
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )
                         , axis.ticks.length = unit(-0.25,"cm")
                         )
  
 
   for (d in sequence(ndevice)){
    h_output = paste0(f_vnam,"_comp.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = secy_path
                    , width    = gg_widthn*0.8
                    , height   = gg_height*0.8
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  
  
}



```


```{r,label="plot-mortality-comp"}

mort_var = mortout_pft[c(1,4:7,11)]


  s_colpfts   = pftinfo$colour
  s_pftlabs   = pftinfo$short
  names(s_pftlabs) = c("1","2","3","4")
  
  mort_labs   = c("BgMort","CstrMort","FireMort","FrezMort","HydroMort","TermMort")
  
  df_now = pft_mortmean %>% filter(mortality %in% mort_var & case_ens %in% survi_ens)
  df_now$pft   = factor(df_now$pft,levels=sequence(npfts))
  df_now$short = rep(mort_labs,each = ncase*nsurvi*npfts)
  gg_now = ggplot(df_now,aes(x = pft, y = value, fill=short))
  gg_now = gg_now + facet_wrap(.~case_ens, scales="free",nrow=2)
  gg_now = gg_now + geom_bar(stat = "identity")
  gg_now = gg_now + scale_x_discrete(labels = s_pftlabs)
  gg_now = gg_now + scale_fill_manual(values = top12)
  gg_now = gg_now + ylab(expression("Mortality (%"~year^-1~")"))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )
                         , axis.ticks.length = unit(-0.25,"cm")
                         )
  
 
   for (d in sequence(ndevice)){
    h_output = paste0("mort-comp.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot     = gg_now
                    , device   = gg_device[d]
                    , path     = tsmort_path
                    , width    = gg_width
                    , height   = gg_height
                    , units    = gg_units
                    , dpi      = gg_depth
                    )
   }
  
  



```




```{r,label="plot-size-structure-time-series"}

dbh_plot  = bydbh                                                        
            
dbh_plot  =  dbh_plot                        %>% 
             mutate(fates_ddbh_canopy = ifelse(is.na(fates_nplant_canopy) | fates_nplant_canopy==0,
                                               fates_ddbh_canopy, fates_ddbh_canopy*ha.2.m2/fates_nplant_canopy),
                    fates_ddbh_ustory = ifelse(is.na(fates_nplant_ustory) | fates_nplant_ustory==0,
                                               fates_ddbh_ustory, fates_ddbh_ustory*ha.2.m2/fates_nplant_ustory),
                    fates_ddbh        = ifelse(is.na(fates_nplant) | fates_nplant==0, 
                                               fates_ddbh, fates_ddbh*ha.2.m2/fates_nplant))


dbh_legend = desc.unit(desc="DBH",unit=untab$cm)
#---~---

dbh_loop = which(fatesvar_new$vnam %in% names(dbh_plot))
gg_dbh   = list()

  for (f in dbh_loop){
  #--- Match variables.
  f_vnam   = fatesvar_new$vnam [f]
  f_desc   = fatesvar_new$desc [f]
  f_unit   = fatesvar_new$unit [f]
  if(f_vnam %in% c("fates_ddbh","fates_ddbh_canopy","fates_ddbh_ustory")){
    f_stack = FALSE
  }else{f_stack  = fatesvar_new$stack[f]}
  
  #f_dbh01  = fatesvar_szpf$dbh01[f]
  f_dbh01  = FALSE #ignore the 0-5cm size class for oak
  cat0("   - ",f_desc,".")
  #---~---
  
  #--- Decide whether to plot the first class.
  if (f_dbh01){
    #--- Keep all classes.
    f_bydbh     = dbh_plot
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs))
    f_coldbhs   = magma(ndbhs,direction=1)
    f_dbhlabs   = dbhinfo$labs
    f_dbhlabs   = parse(text = f_dbhlabs)
    #---~---
  }else{
    #--- Exclude first class.
    bye         = as.numeric(dbh_plot$dbh) %in% 1
    f_bydbh     = dbh_plot[! bye,]
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs)[-1])
    f_coldbhs   = magma(ndbhs,direction=1)[-1]
    f_dbhlabs   = dbhinfo$labs[-1]
    f_dbhlabs   = parse(text = f_dbhlabs)
  }#end if (f_dbh01)
  #---~---
   
  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f_stack){
     gg_now = ggplot(data=f_bydbh,aes_string(x="time",y=f_vnam,group="dbh",fill="dbh"))
     gg_now = gg_now + facet_wrap(.~desc,nrow=5,scales="free")
     gg_now = gg_now + scale_fill_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_area(position=position_stack(reverse = FALSE),show.legend = TRUE)
  }else{
     gg_now = ggplot(data=f_bydbh,aes_string(x="time",y=f_vnam,group="dbh",colour="dbh"))
     gg_now = gg_now + facet_wrap(.~desc,nrow=5,scales="free")
     gg_now = gg_now + scale_colour_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_line(lwd=1.0,show.legend = TRUE)
  }#end if (f_stack)
  #gg_now = gg_now + labs(title=case_desc)
  gg_now = gg_now + scale_x_datetime(date_labels=gg_tfmt)
  gg_now = gg_now + xlab("Simulation time")
  gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz*1.8, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz*1.8
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz*1.8
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f_vnam,"-tsdbh-rev",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tsdbh_path
                  , width    = gg_width*2.5
                  , height   = gg_height*2.5
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg_dbh[[f_vnam]] = gg_now
  #---~---
}


```




```{r,label="plot-annual-ddbh-dynamics"}


year_ddbh  = dbh_plot                        %>%    
             dplyr::select(case,
                           desc,
                           year,
                           dbh,
                           fates_ddbh_canopy,
                           fates_ddbh_ustory,
                           fates_ddbh)         %>% 
             group_by(case,desc,year,dbh)      %>% 
             summarize_all(mean,na.rm=TRUE)    %>% 
             ungroup()                         %>% 
             mutate(year = as.POSIXct(ymd(year, truncated = 2L)))

dbh_loop = which(fatesvar_new$vnam %in% names(year_ddbh))
gg_dbh   = list()

  for (f in dbh_loop){
  #--- Match variables.
  f_vnam   = fatesvar_new$vnam [f]
  f_desc   = fatesvar_new$desc [f]
  f_unit   = fatesvar_new$unit [f]
  f_stack  = fatesvar_new$stack[f]
  f_heat   = TRUE
  #f_dbh01  = fatesvar_szpf$dbh01[f]
  f_dbh01  = FALSE #ignore the 0-5cm size class for oak
  cat0("   - ",f_desc,".")
  #---~---
  
  #--- Decide whether to plot the first class.
  if (f_dbh01){
    #--- Keep all classes.
    f_bydbh     = year_ddbh
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs))
    f_coldbhs   = magma(ndbhs,direction=1)
    f_dbhlabs   = dbhinfo$labs
    f_dbhlabs   = parse(text = f_dbhlabs)
    #---~---
  }else{
    #--- Exclude first class.
    bye         = as.numeric(year_ddbh$dbh) %in% 1
    f_bydbh     = year_ddbh[! bye,]
    f_bydbh$dbh = factor(f_bydbh$dbh,levels=sequence(ndbhs)[-1])
    f_coldbhs   = magma(ndbhs,direction=1)[-1]
    f_dbhlabs   = dbhinfo$labs[-1]
    f_dbhlabs   = parse(text = f_dbhlabs)
  }#end if (f_dbh01)
  #---~---
   
  #--- Initialise plot (decide whether to plot lines or stacks).
  if (f_heat){
     gg_now = ggplot(data=f_bydbh,aes_string(x="year",y="dbh",fill=f_vnam))
     gg_now = gg_now + facet_wrap(.~desc, nrow=2, scales="free")
     gg_now = gg_now + scale_fill_gradient(low="white",high="blue",
                                           name=desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
     gg_now = gg_now + geom_tile(show.legend = TRUE)
  }else{
     gg_now = ggplot(data=f_bydbh,aes_string(x="year",y=f_vnam,group="dbh",colour="dbh"))
     gg_now = gg_now + facet_wrap(.~desc, nrow=5, scales="free")
     gg_now = gg_now + scale_colour_manual(name=dbh_legend,labels=f_dbhlabs,values=f_coldbhs)
     gg_now = gg_now + geom_line(lwd=0.8,show.legend = TRUE)
  }#end if (f_stack)
  #gg_now = gg_now + labs(title=case_desc)
  gg_now = gg_now + scale_x_datetime(date_labels=gg_tfmt)
  gg_now = gg_now + xlab("Simulation time")
  gg_now = gg_now + ylab("Size class (cm)")
  gg_now = gg_now + scale_y_discrete(labels=f_dbhlabs)
  #gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
  #gg_now = gg_now + ylab(bquote(.(f_desc)~"("~cm~plant^-1~yr^-1~")"))
  gg_now = gg_now + theme_bw( base_size = gg_ptsz, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "right"
                         , axis.text.x       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f_vnam,"-yeardbh-scaled",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tsdbh_path
                  , width    = gg_widthn
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg_dbh[[f_vnam]] = gg_now
  #---~---
}#end for (a in age_loop)

```





```{r, label="plot-mortality-by-size"}

mortout_sz = names(bydbh)[grep(pattern="mortality",x=names(bydbh))]
szmort     =   bydbh                                             %>%  
               dplyr::select(all_of(c("case","ens_label",
                                      "case_ens",
                                      "dbh", mortout_sz)))       %>% 
               filter(case_ens %in% survi_ens)

szmort     =      szmort                               %>% 
                  group_by(case,ens_label,
                           case_ens,dbh)               %>% 
                  summarize_at(mortout_sz, 
                               mean, na.rm=TRUE)      %>% 
                  ungroup()
#szmort$case = factor(szmort$case,levels=unique(szmort$case))
szmort$dbh  = factor(szmort$dbh,levels=unique(szmort$dbh))
dbh_labs = dbhinfo$labs
dbh_labs = parse(text = dbh_labs)
mortsz_loop = which(fatesvar_szpf$vnam %in% mortout_sz)
for(m in mortsz_loop){
  m_vnam = fatesvar_szpf$vnam[m] 
  m_desc = fatesvar_szpf$desc[m]
  m_unit = fatesvar_szpf$unit[m]
  m_legend =  1
  cat0("   - ",m_desc,".")
  gg_now = ggplot(szmort, aes_string(x="dbh", y=m_vnam)) 
  gg_now = gg_now + facet_wrap(.~case_ens,nrow=2) 
  gg_now = gg_now + scale_x_discrete(labels=dbh_labs)
  gg_now = gg_now + geom_bar(stat = "identity") 
  gg_now = gg_now + xlab("Size class (cm)")
  gg_now = gg_now + ylab(desc.unit(desc=m_desc,unit=untab[[m_unit]],twolines=TRUE))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz*1.2, 
                                base_family = "Helvetica",
                                base_line_size = 0.5,
                                base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           , angle = 90
                                                           , vjust = 0.5
                                                           , hjust=1
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#

f.output = paste0(m_vnam,"-bydbh",".",gg_device)
dummy = ggsave(     filename = f.output
                  , plot     = gg_now
                  , device   = gg_device
                  , path     = tsmort_path
                  , width    = gg_width
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )

  
}



```




```{r,label="non-statck-PFT-annual-dynamic-plots"}

plot_df  = pft_amean                        %>%  
           filter(case_ens %in% survi_ens)  %>% 
           ungroup()                      %>% 
           mutate(year = as.POSIXct(ymd(year, truncated = 2L)))
plot_df$pft = factor(plot_df$pft,levels=sequence(npfts))


plot_loop = which(fatesvar_new$vnam %in% names(plot_df))
gg_dbh   = list()

  for (f in plot_loop){
  #--- Match variables.
  f_vnam   = fatesvar_new$vnam [f]
  f_desc   = fatesvar_new$desc [f]
  f_unit   = fatesvar_new$unit [f]
  f_colpfts   = pftinfo$colour
  f_pftlabs   = pftinfo$short
  names(f_pftlabs) = c("1","2","3","4")

  cat0("   - ",f_desc,".")
  #---~---
  
  gg_now = ggplot(data=plot_df,aes_string(x="year",y=f_vnam,group="pft",colour="pft"))
  gg_now = gg_now + facet_wrap(.~case_ens,nrow=2)
  gg_now = gg_now + scale_colour_manual(name="",labels=f_pftlabs,values=f_colpfts)
  gg_now = gg_now + geom_line(lwd=1,show.legend = TRUE)
  #gg_now = gg_now + labs(title=case_desc)
  gg_now = gg_now + scale_x_datetime(date_labels=gg_tfmt)
  gg_now = gg_now + xlab("Simulation time")
  gg_now = gg_now + ylab(desc.unit(desc=f_desc,unit=untab[[f_unit]],twolines=TRUE))
 # gg_now = gg_now + ylab(bquote(.(f_desc)~"("~cm~plant^-1~yr^-1~")"))
  gg_now = gg_now + theme_grey( base_size = gg_ptsz*1.2, base_family = "Helvetica",base_line_size = 1.5,base_rect_size =0.5)
  gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#end theme
  #---~---

  #--- Save plot.
  for (d in sequence(ndevice)){
    f.output = paste0(f_vnam,"-annual-unstacked",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tspft_path
                  , width    = gg_width*1.2
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  #---~---

  #--- Write plot settings to the list.
  gg_dbh[[f_vnam]] = gg_now
  #---~---
}#end for (a in age_loop)


```




```{r,label="plot-PFT-size-structure"}

for(n in sequence(ncase)){
  case_now = case_names[n]
  desc_now = case_desc[n]
  nc_conn = nc_open(filename=case_files[n])
  nplant_can = ncvar_get(nc_conn,"FATES_NPLANT_CANOPY_SZPF")
  nplant_uns = ncvar_get(nc_conn,"FATES_NPLANT_USTORY_SZPF")
  nplant = nplant_can + nplant_uns
  nplant = nplant*ha.2.m2   #plant/ha
  dummy  = nc_close(nc_conn)


  np_pine = nplant[3:13,,] #only plants with DBH > 10cm
  np_cedar= nplant[16:26,,]
  np_fir  = nplant[29:39,,]
  np_oak  = nplant[42:52,,]

  np_pine = aperm(np_pine,c(2,3,1))
  np_cedar= aperm(np_cedar,c(2,3,1))
  np_fir  = aperm(np_fir,c(2,3,1))
  np_oak  = aperm(np_oak,c(2,3,1))

  pine_vec = as.vector(np_pine)
  pine_vec = tibble(pine_vec) %>% rename(nplant = pine_vec)
  pine_vec$dbh = rep(dbhinfo$id[3:13],each=ntstamp_each*ninst)
  pine_vec$time= rep(each_tstamp,times=ninst*(ndbhs-2))
  cedar_vec = as.vector(np_cedar)
  cedar_vec = tibble(cedar_vec) %>% rename(nplant = cedar_vec)
  cedar_vec$dbh = rep(dbhinfo$id[3:13],each=ntstamp_each*ninst)
  cedar_vec$time= rep(each_tstamp,times=ninst*(ndbhs-2))
  fir_vec   = as.vector(np_fir)
  fir_vec   = tibble(fir_vec) %>% rename(nplant = fir_vec)
  fir_vec$dbh = rep(dbhinfo$id[3:13],each=ntstamp_each*ninst)
  fir_vec$time= rep(each_tstamp,times=ninst*(ndbhs-2))
  oak_vec   = as.vector(np_oak)
  oak_vec   = tibble(oak_vec) %>% rename(nplant = oak_vec)
  oak_vec$dbh = rep(dbhinfo$id[3:13],each=ntstamp_each*ninst)
  oak_vec$time= rep(each_tstamp,times=ninst*(ndbhs-2))
  np_pft    = rbind(pine_vec,cedar_vec,fir_vec,oak_vec)
  np_pft$pft  = rep(pftinfo$id,each=ntstamp_each*ninst*(ndbhs-2))
  lab         = rep(ens_labels,each=ntstamp_each)
  np_pft$ens_label = rep(lab,times = (ndbhs-2)*npfts)
  
  
  szpf_survi = np_pft                            %>% 
             filter(ens_label %in% survi_lab)    %>% 
             mutate(dbh = ifelse(dbh==3,4,dbh))  %>% 
             group_by(ens_label,pft,dbh,time)    %>% 
             summarize_all(sum,na.rm=TRUE)       %>% 
             ungroup()                           %>% 
             mutate(year=year(time))             %>% 
             filter(year %in% c(lst_s:lst_e))    %>% 
             dplyr::select(-c(time,year))        %>% 
             group_by(ens_label,pft,dbh)         %>% 
             summarize_all(mean)                 %>% 
             ungroup()        
  
  szpf_survi$dbh = factor(szpf_survi$dbh,levels=unique(szpf_survi$dbh))
  dbh_labs     = c(dbhinfo$labs[5:13], "paste(10-20)")
  dbh_labs     = parse(text=dbh_labs)
  names(dbh_labs) = c("5","6","7","8","9","10","11","12","13","4")
   szpf_survi$pft = factor(szpf_survi$pft,levels=unique(szpf_survi$pft))
   s_pftlabs   = pftinfo$short
   s_pftcols   = pftinfo$colour
   names(s_pftlabs) = c("1","2","3","4") 

  plot1 = ggplot(szpf_survi,aes(dbh,nplant)) + geom_bar(stat = "identity") +
       facet_wrap(pft~ens_label,scales="free",labeller=labeller(pft=s_pftlabs),nrow=npfts) + 
       scale_x_discrete(labels = dbh_labs ) +
       labs(title=desc_now,
            y=expression("Number of trees"~"("~plant~ha^-1~")"),
            x="Size class(cm)") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  for (d in sequence(ndevice)){
    f.output = paste0(case_now,"PFT-size-distribution",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = plot1
                  , device   = gg_device[d]
                  , path     = tspft_path
                  , width    = gg_widthn*0.6
                  , height   = gg_height*0.8
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  
  
   plot2 = ggplot(szpf_survi,aes(dbh,nplant,fill=pft)) + geom_bar(stat = "identity") +
       facet_wrap(.~ens_label,scales="free") + 
       scale_x_discrete(labels = dbh_labs ) +
       scale_fill_manual(values=s_pftcols,labels=as_labeller(s_pftlabs))+
       labs(title=desc_now,
            y=expression("Number of trees"~"("~plant~ha^-1~")"),
            x="Size class(cm)") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  for (d in sequence(ndevice)){
    f.output = paste0(case_now,"size-distribution-pftstack",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = plot2
                  , device   = gg_device[d]
                  , path     = tspft_path
                  , width    = gg_widthn*0.4
                  , height   = gg_height*0.5
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
}


```




```{r,label="plot-PFT-size-structure-all13"}

  #tk = c("01","02","03","04","05","06","07","08","09","10","11","12","13","14","15","16")
  #tk = c("17","18","19","20","21","22","23","24","25","26","27","28","29","30","31","32")
  #tk  =c("33","34","35","36")
  

for(n in sequence(ncase)){
  case_now = case_names[n]
  desc_now = case_desc[n]
  nc_conn = nc_open(filename=case_files[n])
  nplant_can = ncvar_get(nc_conn,"FATES_NPLANT_CANOPY_SZPF")
  nplant_uns = ncvar_get(nc_conn,"FATES_NPLANT_USTORY_SZPF")
  nplant = nplant_can + nplant_uns
  nplant = nplant*ha.2.m2   #plant/ha
  dummy  = nc_close(nc_conn)


  np_pine = nplant[1:13,,] 
  np_cedar= nplant[14:26,,]
  np_fir  = nplant[27:39,,]
  np_oak  = nplant[40:52,,]

  np_pine = aperm(np_pine,c(2,3,1))
  np_cedar= aperm(np_cedar,c(2,3,1))
  np_fir  = aperm(np_fir,c(2,3,1))
  np_oak  = aperm(np_oak,c(2,3,1))

  pine_vec = as.vector(np_pine)
  pine_vec = tibble(pine_vec) %>% rename(nplant = pine_vec)
  pine_vec$dbh = rep(dbhinfo$id[1:13],each=ntstamp_each*ninst)
  pine_vec$time= rep(each_tstamp,times=ninst*ndbhs)
  cedar_vec = as.vector(np_cedar)
  cedar_vec = tibble(cedar_vec) %>% rename(nplant = cedar_vec)
  cedar_vec$dbh = rep(dbhinfo$id[1:13],each=ntstamp_each*ninst)
  cedar_vec$time= rep(each_tstamp,times=ninst*ndbhs)
  fir_vec   = as.vector(np_fir)
  fir_vec   = tibble(fir_vec) %>% rename(nplant = fir_vec)
  fir_vec$dbh = rep(dbhinfo$id[1:13],each=ntstamp_each*ninst)
  fir_vec$time= rep(each_tstamp,times=ninst*ndbhs)
  oak_vec   = as.vector(np_oak)
  oak_vec   = tibble(oak_vec) %>% rename(nplant = oak_vec)
  oak_vec$dbh = rep(dbhinfo$id[1:13],each=ntstamp_each*ninst)
  oak_vec$time= rep(each_tstamp,times=ninst*ndbhs)
  np_pft    = rbind(pine_vec,cedar_vec,fir_vec,oak_vec)
  np_pft$pft  = rep(pftinfo$id,each=ntstamp_each*ninst*ndbhs)
  lab         = rep(ens_labels,each=ntstamp_each)
  np_pft$ens_label = rep(lab,times = ndbhs*npfts)
  np_pft$node = substr(np_pft$ens_label,8,8) 
  np_pft$task = substr(np_pft$ens_label,16,17)
  
  
  szpf_survi = np_pft                            %>% 
             filter(ens_label %in% survi_lab)    %>%
             #filter(node==2 & task%in%tk)         %>% 
             mutate(year=year(time))             %>% 
             filter(year %in% c(lst_s:lst_e))    %>% 
             dplyr::select(-c(time,year))        %>% 
             group_by(ens_label,pft,dbh)         %>% 
             summarize_all(mean)                 %>% 
             ungroup()        
  
  szpf_survi$dbh = factor(szpf_survi$dbh,levels=unique(szpf_survi$dbh))
  dbh_labs     = dbhinfo$labs[1:13]
  dbh_labs     = parse(text=dbh_labs)
  names(dbh_labs) = c("1","2","3","4","5","6","7","8","9","10","11","12","13")
   szpf_survi$pft = factor(szpf_survi$pft,levels=unique(szpf_survi$pft))
   s_pftlabs   = pftinfo$short
   s_pftcols   = pftinfo$colour
   names(s_pftlabs) = c("1","2","3","4") 

  plot1 = ggplot(szpf_survi,aes(dbh,nplant)) + geom_bar(stat = "identity") +
       facet_wrap(pft~ens_label,scales="free",labeller=labeller(pft=s_pftlabs),nrow=npfts) + 
       scale_x_discrete(labels = dbh_labs ) +
       labs(title=desc_now,
            y=expression("Number of trees"~"("~plant~ha^-1~")"),
            x="Size class(cm)") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
  
for (d in sequence(ndevice)){
    f.output = paste0(case_now,"PFT-size-distribution-all13",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = plot1
                  , device   = gg_device[d]
                  , path     = tspft_path
                  , width    = gg_widthn*0.6
                  , height   = gg_height*0.8
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  
  
  
   plot2 = ggplot(szpf_survi,aes(dbh,nplant,fill=pft)) + geom_bar(stat = "identity") +
       facet_wrap(.~ens_label,scales="free") + 
       scale_x_discrete(labels = dbh_labs ) +
       scale_fill_manual(values=s_pftcols,labels=as_labeller(s_pftlabs))+
       labs(title=desc_now,
            y=expression("Number of trees"~"("~plant~ha^-1~")"),
            x="Size class(cm)") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  for (d in sequence(ndevice)){
    f.output = paste0(case_now,"size-distribution-pftstack-all13",".",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = plot2
                  , device   = gg_device[d]
                  , path     = tspft_path
                  , width    = gg_width
                  , height   = gg_height
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
}


```



```{r,label="plot-mortality-by-PFT-size-class"}

#selec_month = c("7","8","9","10","11")
mort_which = which(fatesvar_new$vnam %in% mortout_pft)

for(n in sequence(ncase)){
  case_now = case_names[n]
  desc_now = case_desc[n]
  nc_conn = nc_open(filename=case_files[n])
  nplant_can = ncvar_get(nc_conn,"FATES_NPLANT_CANOPY_SZPF")
  nplant_uns = ncvar_get(nc_conn,"FATES_NPLANT_USTORY_SZPF")
  nplant = nplant_can + nplant_uns
  nplant = as.vector(aperm(nplant,c(2,3,1)))
  nplant_ustory = as.vector(aperm(nplant_uns,c(2,3,1)))
  for (m in mort_which){
    m_vnam   = fatesvar_new$vnam [m]
    m_desc   = fatesvar_new$desc [m]
    m_unit   = fatesvar_new$unit [m]
    mort_now = paste0(toupper(m_vnam), "_SZPF")
    mort_df  = ncvar_get(nc_conn,mort_now)
    mort_df  = aperm(mort_df,c(2,3,1))
    mort_df  = tibble(as.vector(mort_df)) 
    names(mort_df) = fatesvar_new$vnam[m]
    mort_df$dbh = rep(index_scpf$scls,each = ntstamp_each*ninst)
    mort_df$pft = rep(index_scpf$pft,each = ntstamp_each*ninst)
    mort_df$nplant = nplant
    mort_df$nplant_ustory = nplant_ustory
    mort_df$time   = rep(each_tstamp,times=ninst*ndbhs*npfts)
    lab            = rep(ens_labels,each=ntstamp_each)
    mort_df$ens_label = rep(lab,times = ndbhs*npfts)
    mort_pfsz = mort_df                             %>% 
                filter(ens_label %in% survi_lab)    %>% 
                mutate_at(m_vnam, 
                          ~ find_mort(x=.x,n= if_else(.x=="fates_mortality_ustory", 
                                                      .data$nplant_ustory, 
                                                      .data$nplant))) %>% 
               #mutate(month = month(time))                            %>% 
               #filter(month %in% selec_month)                         %>% 
               group_by(ens_label,pft,dbh)                            %>% 
               select(-c(time,nplant_ustory,nplant))                  %>% 
               summarize_all(mean,na.rm=TRUE)                         %>% 
               ungroup()    
  #size_pft = rbind(sz_20,size_pft)


  mort_pfsz$pft = factor(mort_pfsz$pft,levels=unique(mort_pfsz$pft))
  mort_pfsz$dbh = factor(mort_pfsz$dbh,levels=unique(mort_pfsz$dbh))
  pft_labs     = pftinfo$short
  names(pft_labs)=c("1","2","3","4")
  dbh_labs     = dbhinfo$labs[1:13]
  dbh_labs     = parse(text=dbh_labs)
  names(dbh_labs) = c("1","2","3","4","5","6","7","8","9","10","11","12","13")

  plot = ggplot(mort_pfsz,aes_string("dbh",m_vnam)) + geom_bar(stat = "identity") +
       facet_wrap(pft~ens_label, labeller=labeller(pft=pft_labs)) + 
       scale_x_discrete(labels = dbh_labs ) +
       labs(title=desc_now,
            y=desc.unit(desc=m_desc,unit=untab[[m_unit]],twolines=TRUE),
            x="Size class(cm)") +
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  for (d in sequence(ndevice)){
    f.output = paste0(case_now,m_vnam,"-pfsz-zero.",gg_device[d])
    dummy = ggsave( filename = f.output
                  , plot     = plot
                  , device   = gg_device[d]
                  , path     = tsmort_path
                  , width    = gg_widthn*0.6
                  , height   = gg_height*0.6
                  , units    = gg_units
                  , dpi      = gg_depth
                  )#end ggsave
  }#end for (o in sequence(nout))
  }
  
  dummy  = nc_close(nc_conn)


  
}


```




```{r,label="plot-monthly-mean-fire-variables"}

fire_vars = c("fates_burnfrac","fates_fdi","fates_ros",
              "fates_fuel_amount","fates_fuel_bulkd","fates_fire_intensity",
              "fates_rxfire_burnfrac",
              "fates_fire_intensity_burnfrac","fates_rxfire_intensity_burnfrac")
fire_df    = hlm1d                                                 %>%
             dplyr::select(all_of(c("case","ens_label",
                                    "time",fire_vars)))            %>% 
             mutate(#fates_fire_intensity_burnfrac = ifelse(fates_burnfrac == 0, 0,
                    #      fates_fire_intensity_burnfrac / (fates_burnfrac/(day.sec*yr.day))),
                   #fates_rxfire_intensity_burnfrac = ifelse(fates_rxfire_burnfrac == 0, 0,
                   #       fates_rxfire_intensity_burnfrac / (fates_rxfire_burnfrac/(day.sec*yr.day))),
                   case_ens  = paste0(case,"_",ens_label))

fire_cmean = fire_df                                               %>%
             mutate(month = month(time),
                    year  = year(time))                            %>% 
             filter(year %in% c(lst_s:lst_e))                      %>% 
             group_by(case,ens_label,case_ens,month)               %>% 
             dplyr::select(-c(time,year))                          %>% 
             summarize_all(mean,na.rm=TRUE)                        %>% 
             ungroup()                                             %>% 
             filter(case_ens %in% survi_ens)

fire_qupr  = fire_df                                               %>% 
             mutate(month = month(time),
                    year  = year(time))                            %>% 
             filter(year %in% c(lst_s:lst_e))                      %>% 
             group_by(case,ens_label,case_ens,month)               %>%
             dplyr::select(-c(time,year))                          %>% 
             summarize_all(quantile,probs=qupr_ribbon,names=FALSE
                          , na.rm=TRUE)                            %>% 
             ungroup()                                             %>% 
             rename_at(all_of(fire_vars),
                       function(x) paste0(x,"_qupr"))              %>% 
             filter(case_ens %in% survi_ens)

fire_qlwr  = fire_df                                               %>% 
             mutate(month = month(time),
                    year  = year(time))                            %>% 
             filter(year %in% c(lst_s:lst_e))                      %>%
             group_by(case,ens_label,case_ens,month)               %>%
             dplyr::select(-c(time,year))                          %>% 
             summarize_all(quantile,probs=qlwr_ribbon,names=FALSE
                          , na.rm=TRUE)                            %>% 
             ungroup()                                             %>% 
             rename_at(all_of(fire_vars),
                       function(x) paste0(x,"_qlwr"))              %>% 
             filter(case_ens %in% survi_ens)

fire_cmean  = merge(x=fire_cmean,y=merge(x=fire_qupr,y=fire_qlwr,
                                         by=c("case","ens_label","case_ens","month")),
                    by=c("case","ens_label","case_ens","month"))
fire_cmean = fire_cmean %>% arrange(case,ens_label,case_ens,month)

## observed seasonal variations in burned area of lightning-caused fires in California
#layers      = st_layers(dsn=frap_raw) 
#wldfire     = st_read(frap_raw, layer = layers$name[1])
#wldfire     = wldfire                                    %>% 
#              filter(CAUSE==1)                           %>% 
#              dplyr::select(YEAR_,ALARM_DATE,Shape_Area) %>% 
#              mutate(month=month(ALARM_DATE)
#                    ,case=0)                             %>% 
#              as.data.frame()                            %>% 
#              dplyr::select(-Shape)                      %>% 
#              group_by(case,month)                       %>% 
#              summarize(fates_burnfrac = mean(Shape_Area,na.rm=TRUE)
#                       ,fates_burnfrac_qupr = quantile(Shape_Area,probs=qupr_ribbon,names=FALSE, na.rm=TRUE)
#                       ,fates_burnfrac_qlwr = quantile(Shape_Area,probs=qlwr_ribbon,names=FALSE, na.rm=TRUE)) #%>% 
#              filter(!is.na(month)) %>% ungroup()
#ba_df      = fire_cmean %>% dplyr::select(case,month,fates_burnfrac,fates_burnfrac_qupr,fates_burnfrac_qlwr) 
#ba_df      = rbind(wldfire,ba_df) %>% arrange(case,month)

byfuel$fuel = factor(byfuel$fuel,levels=sequence(nfuels))
byfuel      = byfuel %>% mutate(case_ens=paste0(case,"_",ens_label))
fuelmmean_bysize = byfuel                                          %>% 
              mutate(month = month(time),
                     year  = year(time))                           %>% 
              filter(year %in% c(lst_s:lst_e) & 
                       case_ens %in% survi_ens)                    %>%
              group_by(case,ens_label,case_ens,fuel,month)         %>% 
              dplyr::select(-c(time,year))                         %>% 
              summarize_all(mean,na.rm=TRUE)                       %>% 
              ungroup()

fuelamean_bysize = byfuel                                          %>% 
              mutate(year = year(time))                            %>% 
              filter(case_ens %in% survi_ens)                      %>% 
              group_by(case,ens_label,case_ens,fuel,year)          %>% 
              dplyr::select(-time)                                 %>% 
              summarize_all(mean,na.rm=TRUE)                       %>% 
              ungroup()

sumfuel          = byfuel                                          %>%
                   filter(case_ens %in% survi_ens &
                          fuel !="4")                              %>% 
                   group_by(case,ens_label,case_ens,time)          %>% 
                   dplyr::select(-fuel)                            %>% 
                   summarize_all(sum,na.rm=TRUE)                   %>% 
                   ungroup()                                       %>% 
                   mutate(month = month(time),
                          year  = year(time))                      %>% 
                   filter(year %in% c(lst_s:lst_e))                %>%
                   group_by(case,ens_label,case_ens,month)         %>% 
                   dplyr::select(-c(time,year))                    %>% 
                   summarize_all(mean,na.rm=TRUE)                  %>% 
                   ungroup()                                     

#sumfuel          = sumfuel %>% left_join(fire_mmean,by=c("case","month"))

fuelmmean_plot = ggplot(fuelmmean_bysize, aes(x=month,y=fates_fuel_amount,group=fuel,color=fuel)) +
  geom_line(lwd=0.5) +
  scale_x_continuous( breaks = sequence(12)) +
  scale_color_manual(labels = fuelinfo$short,values=fuelinfo$colour)+
  facet_wrap(case~ens_label,labeller=labeller(case=case_desc),nrow=2)  +
  labs(x="Month",y="Fuel amount")
ggsave(fuelmmean_plot,path=plot_main,filename="fueload-mmean.jpeg",width=gg_widthn*0.6,height=gg_height*0.5,dpi=gg_depth)  

sfuelmmean_plot = ggplot(sumfuel, aes(x=month,y=fates_fuel_amount)) +
  geom_line(lwd=0.5) +
  scale_x_continuous(breaks = sequence(12)) +
  facet_wrap(case~ens_label,labeller=labeller(case=case_desc),nrow=2)  +
  labs(x="Month",y="Total fuel amount omits trunk")
ggsave(sfuelmmean_plot,path=plot_main,filename="sumfueload-mmean.jpeg",width=gg_widthn*0.6,height=gg_height*0.5,dpi=gg_depth)  

fuelamean_plot = ggplot(fuelamean_bysize, aes(x=year,y=fates_fuel_amount,group=fuel,color=fuel)) + geom_line(lwd=1) +
  #scale_x_discrete(breaks=1:12) +
  scale_color_manual(labels = fuelinfo$short,values=fuelinfo$colour)+
  facet_wrap(case~ens_label,labeller=labeller(case=case_desc),nrow=2)  +
  labs(x="",y="Fuel amount")
ggsave(fuelamean_plot,path=plot_main,filename="fueload-amean.jpeg",width=gg_widthn,height=gg_height*0.5,dpi=gg_depth)  


             
plot_fire = c("fates_burnfrac","fates_rxfire_burnfrac","fates_fire_intensity", "fates_rxfire_intensity_burnfrac","fates_fire_intensity_burnfrac")
fire_loop = which(hlm1dvar$vnam %in% fire_vars)
gg_fire   = list()
tf        = 2e+10
for(h in fire_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_vqlwr  = paste0(h_vnam,"_qlwr")
   h_vqupr  = paste0(h_vnam,"_qupr")
   h_legend = v == 1
   cat0("   - ",h_desc,".")

   fire_now        = fire_cmean
   
   gg_now = ggplot(fire_now,aes(x=month))
   gg_now = gg_now + facet_wrap(case~ens_label,labeller=labeller(case=case_desc),nrow=2)
   gg_now = gg_now + geom_line(aes_string(y=h_vnam)
                              ,lwd=0.5,alpha=1,show.legend=TRUE)
   #gg_now = gg_now + geom_line(data=wldfire,aes(y=fates_burnfrac / tf),
                               #show.legend=FALSE,color="black")
   #gg_now = gg_now + scale_colour_manual(name="",aesthetics="colour",values=top12,labels=case_desc)
   gg_now = gg_now + geom_ribbon( data = fire_now 
                                , aes_string(x = "month",ymin=h_vqlwr,ymax=h_vqupr)
                                , alpha       = alpha_ribbon*0.5
                                , show.legend = h_legend
                                )
   #gg_now = gg_now + scale_fill_manual(name="",values=top12,labels=case_desc)

   gg_now = gg_now + scale_y_continuous(name=desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))#,
                                        #sec.axis = sec_axis(~.*tf,name=expression("FRAP"~"["*m^2*"]")))
   
   #gg_now = gg_now + labs(title=h_desc)
   gg_now = gg_now + scale_x_continuous( breaks = sequence(12))#end scale_x_continuous
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 1.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,"_monthlymean_Survi.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot      = gg_now
                    , device    = gg_device[d]
                    , path      = secy_path
                    , width     = gg_widthn*0.7
                    , height    = gg_height*0.7
                    , units     = gg_units
                    , dpi       = gg_depth
                    )
   }
  gg_fire[[h_vnam]] = gg_now      

}


# If sought, plot images on screen
if (gg_screen) gg_fire


```





```{r,label="annual-fire-dynamic",message=FALSE,results='hide'}

fire_month = c("6","7","8","9","10")
plot_fire = fire_vars[c(1,4,5,7,8)] 

fiream_df =  fire_df                                             %>% 
                  dplyr::select(all_of(c("case","ens_label",
                                         "case_ens","time",plot_fire)))
                 
fiream_df = fiream_df                                              %>% 
                 mutate(month=month(time),year=year(time))         %>% 
                 filter(#year %in% c(1769:1799),
                        month %in% fire_month,
                        case_ens %in% survi_ens)                   %>% 
                 group_by(case,ens_label,case_ens,year)            %>%
                 summarize_all(mean,na.rm=TRUE)                    %>% 
                 ungroup()                                        # %>% 
                # left_join(cags_quantyr,by="year")
             
fire_loop = which(hlm1dvar$vnam %in% plot_fire)
gg_fire   = list()
for(h in fire_loop){
   h_vnam   = hlm1dvar$vnam[h] 
   h_desc   = hlm1dvar$desc [h]
   h_short  = hlm1dvar$short[h]
   h_unit   = hlm1dvar$unit [h]
   h_legend = 1
   cat0("   - ",h_desc,".")

   fire_now        = fiream_df
  
   gg_now = ggplot(data=fire_now)
   gg_now = gg_now + facet_wrap(case~ens_label,labeller=labeller(case=case_desc),nrow=2)
   gg_now = gg_now + geom_line(aes_string(x="year",y=h_vnam),lwd=0.5,alpha=0.8,show.legend=TRUE)
   #gg_now = gg_now + geom_line(aes(x=year,y=yr_mean,color="FRAP"))
   #gg_now = gg_now + scale_color_manual(labels=as_labeller(case_desc),values=c(ens_colour,top12[1]))
   #gg_now = gg_now + scale_colour_manual(name=""
   #                                     ,values=c("Model"= ens_colour[1]
   #                                              ,"FRAP" = site_colour))
  # gg_now = gg_now + geom_ribbon( aes(x = year,ymin=yr_qlwr,ymax=yr_qupr)
  #                              , alpha       = alpha_ribbon
  #                              , inherit.aes = FALSE
  #                              , show.legend = h_legend
  #                              , colour      = "transparent"
  #                              )
   #gg_now = gg_now + scale_y_log10()
   #gg_now = gg_now + scale_x_continuous( breaks = sequence(12))#end scale_x_continuous
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_short,unit=untab[[h_unit]],dxpr=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz*1.2, base_family = "Helvetica",base_line_size= 0.5,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz*1.2)
                         , axis.text.x        = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,"_annual.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot      = gg_now
                    , device    = gg_device[d]
                    , path      = secy_path
                    , width     = gg_widthn
                    , height    = gg_height*0.7
                    , units     = gg_units
                    , dpi       = gg_depth
                    )
   }
  gg_fire[[h_vnam]] = gg_now      

}


# If sought, plot images on screen
if (gg_screen) gg_fire




```




```{r,label="size-distribution"}

dbh_labs = c(dbhinfo$labs[5:13], "paste(10-20)")
dbh_labs = parse(text=dbh_labs)
names(dbh_labs) = c(dbhinfo$id[5:13],"4")

sz_nplant = bydbh %>% dplyr::select(case, ens_label, time, dbh, fates_nplant) %>% filter(dbh%in%dbhinfo$id[3:13] & ens_label %in% survi_lab) 
sz_nplant = sz_nplant                          %>% 
            mutate(dbh=ifelse(dbh==3,4,dbh),
                   case_ens = paste0(case,"_",
                                 ens_label))   %>% 
            group_by(case,ens_label,
                     case_ens,time,dbh)        %>% 
            summarize_all(sum,na.rm=TRUE)      %>% 
            ungroup()                          %>% 
            mutate(year=year(time))            %>% 
            filter(year %in% c(lst_s:lst_e))
            
sz_nplant = sz_nplant %>% group_by(case,ens_label,case_ens,dbh) %>% 
            summarize(fates_nplant = mean(fates_nplant)) %>% ungroup()
sz_nplant = sz_nplant %>%  arrange(case,ens_label,case_ens,dbh)
sz_nplant$dbh = factor(sz_nplant$dbh,levels=unique(sz_nplant$dbh))

gg_now = ggplot(sz_nplant, aes(dbh, fates_nplant)) 
gg_now = gg_now + facet_wrap(case~ens_label,
                             labeller=labeller(case=case_desc),
                             scales="free",nrow=2) 
gg_now = gg_now + scale_x_discrete(labels=dbh_labs)
gg_now = gg_now + geom_bar(stat = "identity") 
gg_now = gg_now + xlab("Size class (cm)")
gg_now = gg_now + ylab(expression("( Plant"~ha^-2~")"))
gg_now = gg_now + theme_grey( base_size = gg_ptsz*1.2, base_family = "Helvetica",base_line_size = 0.5,base_rect_size =0.5)
gg_now = gg_now + theme( legend.position   = "bottom"
                         , axis.text.x       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           , angle = 90
                                                           , vjust = 0.5
                                                           , hjust=1
                                                           )#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz*1.2
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         )#

f.output = paste0("NPALNT","-bydbh-",".",gg_device[1])
dummy = ggsave(     filename = f.output
                  , plot     = gg_now
                  , device   = gg_device[d]
                  , path     = tsdbh_path
                  , width    = gg_widthn*1.2
                  , height   = gg_height*1.2
                  , units    = gg_units
                  , dpi      = gg_depth
                  )

```





```{r,label="plot-fuel-by-patch"}

fuel_age   = byage                                                     %>% 
             dplyr::select(case,ens_label,time,
                           age,fates_fuel_amount)                      %>% 
             mutate(year = year(time))                                 %>% 
             filter(ens_label %in% survi_lab)                          %>% 
             group_by(case,ens_label,year,age)                         %>% 
             summarize(fates_fuel_amount = mean(fates_fuel_amount))    %>% 
             ungroup()
             
plot_fire = "fates_fuel_amount" 
fire_loop = which(fatesvar_new$vnam %in% plot_fire)
gg_fire   = list()
for(h in fire_loop){
   h_vnam   = fatesvar_new$vnam[h] 
   h_desc   = fatesvar_new$desc [h]
   h_unit   = fatesvar_new$unit [h]
   h_label  = ageinfo$key
   h_legend = v == 1
   cat0("   - ",h_desc,".")

   fire_now        = fuel_age
   fire_now$age   = factor(fire_now$age,levels=unique(fire_now$age))
   
   gg_now = ggplot(data=fire_now)
   gg_now = gg_now + geom_line(aes_string(x="year",y=h_vnam,group="age",colour="age"),lwd=1,alpha=0.8,show.legend=TRUE)
   gg_now = gg_now + facet_wrap(case~ens_label,nrow=2,labeller=labeller(case=case_desc))
   gg_now = gg_now + scale_colour_manual(values=ageinfo$colour,labels = h_label)
   gg_now = gg_now + labs(title=h_desc)
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_desc,unit=untab[[h_unit]],twolines=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz, base_family = "Helvetica",base_line_size= 1,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz)
                         , axis.text.x        = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm"))#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,"_bypatch.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot      = gg_now
                    , device    = gg_device[d]
                    , path      = tsage_path
                    , width     = gg_width
                    , height    = gg_height
                    , units     = gg_units
                    , dpi       = gg_depth
                    )
   }
  gg_fire[[h_vnam]] = gg_now      

}


# If sought, plot images on screen
if (gg_screen) gg_fire
            



```




```{r,label="plot-agb-by-patch-pft"}

scale_age = function(x,p){
  scale = x*p
  return(scale)
}
plot_age = nc_byage[5]
age_pref  = gsub(pattern="_AP$",replacement="",x=plot_age)
age_pref  = gsub(pattern="_APFC$",replacement="",x=plot_age)
age_pref  = gsub(pattern="_APPF$",replacement="",x=plot_age)
age_pref  = tolower(age_pref)
agevar_idx= match(age_pref,fatesvar_ap$vnam)
gg_age   = list()
for(a in sequence(length(plot_age))){
   h        = agevar_idx[a]
   fvar_now = plot_age[a]
   h_vnam   = fatesvar_ap$vnam[h] 
   h_desc   = fatesvar_ap$desc [h]
   h_unit   = fatesvar_ap$unit [h]
   h_label  = ageinfo$key
   h_legend = v == 1
   cat0("   - ",h_desc,".")
   age_df  = NULL
   for(n in sequence(ncase)){
   case_now = case_names[n]
   desc_now = case_desc[n]
   nc_conn  = nc_open(filename=case_files[n])
   vegc     = ncvar_get(nc_conn,fvar_now)
   parea    = ncvar_get(nc_conn,"FATES_PATCHAREA_AP")
   vegc     = aperm(vegc,c(2,3,1))
   vegc     = as.vector(vegc)
   parea    = aperm(parea,c(2,3,1))
   parea    = as.vector(parea)
   parea    = rep(parea,times = npfts)
   vegc     = tibble(vegc)
 
   names(vegc) = h_vnam
   vegc$parea  = parea
   vegc$desc = desc_now
   lab = rep(ens_labels,each = ntstamp_each)
   vegc$ens_label = rep(lab,times=nages*npfts)
   vegc$pft       = rep(pftinfo$id, each = ninst*ntstamp_each*nages)
   agelab         = rep(ageinfo$key,each = ninst*ntstamp_each)
   vegc$age       = rep(agelab, times=npfts)
   vegc$time      = rep(each_tstamp, times = ninst*nages*npfts)
   age_df         = rbind(age_df,vegc)
   dummy  = nc_close(nc_conn)
   rm(vegc)}
   
   age_df   = age_df                       %>% 
              mutate_at(h_vnam,
                        ~ scale_age(x=.x,p=.data$parea)) %>%
              mutate_at(h_vnam,
                        ~ exclude_zero(var=.x))  %>% 
              mutate(year = year(time))    %>% 
              filter(ens_label %in% survi_lab) %>% 
                    # year %in% c(1789:1799)) %>% 
              group_by(desc,ens_label,pft,
                       age)                %>% 
              dplyr::select(-c(time,year)) %>% 
              summarize_all(mean,
                            na.rm=TRUE)    %>% 
              ungroup()                    
   age_df$pft = factor(age_df$pft,levels=unique(age_df$pft))
   s_pftlabs   = pftinfo$short
   names(s_pftlabs) = c("1","2","3","4")   
   gg_now = ggplot(data=age_df,aes_string(x="age",y=h_vnam,fill="pft"))
   gg_now = gg_now + geom_bar(stat = "identity")
   gg_now = gg_now + facet_wrap(desc~ens_label,nrow=ncase,scales="free")
   gg_now = gg_now + scale_fill_manual(values=pftinfo$colour,labels=as_labeller(s_pftlabs))
   gg_now = gg_now + xlab(element_blank())
   gg_now = gg_now + ylab(desc.unit(desc=h_desc,unit=untab[[h_unit]],twolines=TRUE))
   gg_now = gg_now + theme_grey( base_size = gg_ptsz*1.5, base_family = "Helvetica",base_line_size= 1,base_rect_size =0.5)
   gg_now = gg_now + theme( legend.position   = "right"
                         , legend.text=element_text(size=gg_ptsz*1.5)
                         , axis.text.x        = element_text( size   = gg_ptsz*1.5
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           , angle = 90
                                                           , vjust = 0.5
                                                           , hjust=1)#end element_text
                         , axis.text.y       = element_text( size   = gg_ptsz*1.5
                                                           , margin = unit(rep(0.35,times=4),"cm")
                                                           )#end element_text
                         , axis.ticks.length = unit(-0.25,"cm")
                         ) #end theme
  
 
   # Save plots.
   for (d in sequence(ndevice)){
    h_output = paste0(h_vnam,"_bypatch.",gg_device[d])
    dummy    = ggsave( filename = h_output
                    , plot      = gg_now
                    , device    = gg_device[d]
                    , path      = tsage_path
                    , width     = gg_width*1.5
                    , height    = gg_height*1.2
                    , units     = gg_units
                    , dpi       = gg_depth
                    )
   }
  gg_age[[h_vnam]] = gg_now      

}


# If sought, plot images on screen
if (gg_screen) gg_age
            



```